---
title: "Insulin Clamp Study on Mck-TSC1 Mice"
author: "Dave Bridges and Nathan Qi"
date: "March 21, 2012"
output:
  pdf_document:
    highlight: tango
    keep_tex: yes
    number_sections: yes
    toc: yes
  html_document:
    highlight: tango
    keep_md: yes
    number_sections: yes
    toc: yes
---

```{r global_options, include=FALSE}
library(knitr)
library(tidyr)
#figures makde will go to directory called figures, will make them as both png and pdf files 
opts_chunk$set(fig.path='figures/',
               echo=FALSE, warning=FALSE, message=FALSE,dev=c('png','pdf'))
options(scipen = 2, digits = 3)
# set echo and message to TRUE if you want to display code blocks and code output respectively

knitr::knit_hooks$set(inline = function(x) {
  knitr:::format_sci(x, 'md')
})


superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
  arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
  length = length, ...)

  
se <- function(x) sd(x, na.rm=T)/sqrt(length(x))

#load these packages, nearly always needed
library(tidyr)
library(dplyr)

# sets maize and blue color scheme
color.scheme <- c('#00274c', '#ffcb05')
```

# Experiment Summary

```{r data files}
data_file <- "clamp summary.csv"
data <- read.csv(data_file)

kable(data[order(data$Genotype),], caption = "All Data")
```

## Data Summary

```{r data-summary}
library(dplyr)
se <- function(x) sd(x, na.rm=T)/sqrt(length(x))
summary.data <- 
  data %>%
  mutate(Genotype = relevel(Genotype,ref="Wild-Type")) %>%
  group_by(Genotype) %>%
  summarize_at(.vars=vars(Basal.Glucose:Heart),.funs=funs(mean(.,na.rm=T),se))
```

The mean data is compiled with the standard errors

## Statistics

```{r data-analysis}
Clamp.GIR.fit <- t.test(Clamp.GIR~Genotype, data=data)
Clamp.AUC.fit <- t.test(Clamp.AUC~Genotype, data=data)
Basal.HGP.fit <- t.test(Basal.HGP~Genotype, data=data)
Clamp.HGP.fit <- t.test(Clamp.HGP~Genotype, data=data)
sHGP.fit <- t.test(Clamp.sHGP~Genotype, data=data)
Clearance.Basal.fit <- t.test(Clearance.Basal~Genotype, data=data)
Clearance.Clamp.fit <- t.test(Clearance.Clamp~Genotype, data=data)
Insulin.fit <- t.test(Insulin.Basal~Genotype, data=data)
Gastroc.fit <- t.test(Gastroc~Genotype, data=data)
V.fat.fit <- t.test(V.fat~Genotype, data=data)
S.fat.fit <- t.test(S.fat~Genotype, data=data)
BAT.fit <- t.test(BAT~Genotype, data=data)
Heart.fit <- t.test(Heart~Genotype, data=data)
t.tests <- data.frame(Clamp.GIR = Clamp.GIR.fit$p.value,
                      Clamp.AUC = Clamp.AUC.fit$p.value,
                      Basal.HGP = Basal.HGP.fit$p.value,
                      Clamp.HGP = Clamp.HGP.fit$p.value,
                      sHGP = sHGP.fit$p.value,
                      Clearance.Basal = Clearance.Basal.fit$p.value,
                      Clearance.Clamp = Clearance.Clamp.fit$p.value,
                      Insulin = Insulin.fit$p.value,
                      Gastroc =  Gastroc.fit$p.value,
                      V.fat = V.fat.fit$p.value,
                      S.fat = S.fat.fit$p.value,
                      BAT = BAT.fit$p.value,
                      Heart.fit = Heart.fit$p.value)
```


```{r avg-tables}
avg.table <- 
  summary.data %>% select(ends_with('mean')) %>% t

library(tibble)
avg.table <-
  as.tibble(avg.table) %>%
  mutate(Measurement = rownames(avg.table)) 

colnames(avg.table) <- c('WT',"KO","Measurement")
avg.table %>%
  mutate(Difference = KO-WT,
         Pct.Difference = Difference/WT*100) %>% 
  filter(WT >0) %>%
  kable(caption="Summary statistics for each measurement")
```

```{r t-tests}
kable(t(t.tests), caption = "T-Tests, Unadjusted")
kable(t(p.adjust(t.tests)) , caption = "Nominal and Benjamini-Hochberg Adjusted T-Tests")
```


# Graphs

## Glucose Infusion Rate

```{r gir}
plot <- barplot(summary.data$Clamp.GIR_mean, 
                names.arg=(summary.data$Genotype),  
                ylab="Glucose Infusion Rate",
                main="Glucose Infusion Rate (mg/kg/min)",
                ylim=c(0,32),
                las=1)
superpose.eb(plot, summary.data$Clamp.GIR_mean, summary.data$Clamp.GIR_se)
```

## Glucose Turnover

```{r glucose-turnover}
plot.data <-
  summary.data %>%
  select(Genotype, contains('Gtr'))

plot <- barplot(as.matrix(plot.data[2:3]), 
                beside=T,  
                ylab="Glucose Turnover (mg/kg/min)",
                main="Glucose Turnover Rate",
                names.arg=c("Basal","Insulin"),
                ylim=c(0,60),
                las=1)
superpose.eb(plot, as.matrix(plot.data[2:3]), as.matrix(plot.data[4:5]))
legend("topleft", levels(summary.data$Genotype), bty="n",fill=grey.colors(2),)
```

## Endogenous Glucose Production

```{r glucose-production}
plot.data <-
  summary.data %>%
  select(Genotype, contains('.HGP'))

plot <- barplot(as.matrix(plot.data[2:3]), 
                beside=T,  
                ylab="Glucose Production (mg/kg/min)",
                main="Endogenous \nGlucose Production",
                names.arg=c("Basal","Insulin"),
                ylim=c(0,60),
                las=1)
superpose.eb(plot, as.matrix(plot.data[2:3]), as.matrix(plot.data[4:5]))
legend("topleft", levels(summary.data$Genotype), bty="n",fill=grey.colors(2),)
```

## Clamp Summary

```{r clamp-summary}
par(mfrow=c(1,3))

plot <- barplot(summary.data$Clamp.GIR_mean, 
                names.arg=(summary.data$Genotype),  
                ylab="Glucose Infusion Rate (mg/kg/min)",
                main="Glucose Infusion Rate \nDuring Clamp",
                ylim=c(0,32),
                las=1,
                col=grey.colors(2))
superpose.eb(plot, summary.data$Clamp.GIR_mean, summary.data$Clamp.GIR_se)

plot.data <-
  summary.data %>%
  select(Genotype, contains('Gtr'))

plot <- barplot(as.matrix(plot.data[2:3]), 
                beside=T,  
                ylab="Glucose Turnover (mg/kg/min)",
                main="Glucose Turnover Rate",
                names.arg=c("Basal","Insulin"),
                ylim=c(0,60),
                las=1)
superpose.eb(plot, as.matrix(plot.data[2:3]), as.matrix(plot.data[4:5]))
legend("topleft", levels(summary.data$Genotype), bty="n",fill=grey.colors(2),)

plot.data <-
  summary.data %>%
  select(Genotype, contains('.HGP'))

plot <- barplot(as.matrix(plot.data[2:3]), 
                beside=T,  
                ylab="Glucose Production (mg/kg/min)",
                main="Endogenous \nGlucose Production",
                names.arg=c("Basal","Insulin"),
                ylim=c(0,40),
                las=1)
superpose.eb(plot, as.matrix(plot.data[2:3]), as.matrix(plot.data[4:5]))
legend("topleft", levels(summary.data$Genotype), bty="n",fill=grey.colors(2),)
```

### Suppression of Endogenous Glucose Production

```{r sHGP}

plot <- barplot(summary.data$Clamp.sHGP_mean, 
                names.arg=(summary.data$Genotype),  
                ylab="Percent Suppression by Insulin",
                main="Suppression of Endogenous Glucose Production",
                ylim=c(0,100),
                las=1)
superpose.eb(plot, summary.data$Clamp.sHGP_mean, summary.data$Clamp.sHGP_se)
```

## Insulin Levels

```{r insulin}
plot.data <-
  summary.data %>%
  select(Genotype, contains('Insuli'))

plot <- barplot(as.matrix(plot.data[2:3]), 
                beside=T,  
                ylab="Insulin Levels (uU/mL)",
                main="Insulin During Clamp",
                names.arg=c("Basal","Insulin"),
                ylim=c(0,25),
                las=1)
superpose.eb(plot, as.matrix(plot.data[2:3]), as.matrix(plot.data[4:5]))
legend("topleft", levels(summary.data$Genotype), bty="n",fill=grey.colors(2),)
```

# Glucose Clearance Rate

Blood glucose clearance

```{r glucose-clearance}
plot.data <-
  summary.data %>%
  select(Genotype, contains('Clearance'))

plot <- barplot(as.matrix(plot.data[2:3]), 
                beside=T,  
                ylab="Glucose Clearance (mL/kg/min)",
                main="Glucose Clearance Rate ",
                names.arg=c("Basal","Insulin"),
                ylim=c(0,40),
                las=1)
superpose.eb(plot, as.matrix(plot.data[2:3]), as.matrix(plot.data[4:5]))
legend("topleft", levels(summary.data$Genotype), bty="n",fill=grey.colors(2),)
```

# Tissue Glucose Clearance

## Gasctrocnemius Gluose Uptake During Clamp

```{r gastroc}

plot <- barplot(summary.data$Gastroc_mean, 
                names.arg=(summary.data$Genotype),  
                ylab="2-Deoxyglucose Uptake (ug/g/min)",
                main="Gastrocnemius",
                ylim=c(0,17),
                las=1)
superpose.eb(plot, summary.data$Gastroc_mean, summary.data$Gastroc_se)
```

## Visceral Fat Gluose Uptake During Clamp

```{r v-fat}
plot <- barplot(summary.data$V.fat_mean, 
                names.arg=summary.data$Genotype,  
                ylab="2-Deoxyglucose Uptake (ug/g/min)",
                main="Visceral Adipose",
                ylim=c(0,8),
                las=1)
superpose.eb(plot, summary.data$V.fat_mean, summary.data$V.fat_se)
```

## Subcutaneous Fat Gluose Uptake During Clamp

```{r s-fat}
plot <- barplot(summary.data$S.fat_mean, 
                names.arg=summary.data$Genotype,  
                ylab="2-Deoxyglucose Uptake (ug/g/min)",
                main="Subcutaneous Adipose",
                ylim=c(0,30),
                las=1)
superpose.eb(plot, summary.data$S.fat_mean, summary.data$S.fat_se)
```

## Heart Gluose Uptake During Clamp

```{r heart}
plot <- barplot(summary.data$Heart_mean, 
                names.arg=summary.data$Genotype,  
                ylab="2-Deoxyglucose Uptake (ug/g/min)",
                main="Heart",
                ylim=c(0,550),
                las=1)
superpose.eb(plot, summary.data$Heart_mean, summary.data$Heart_se)
```

## Brown Adipose Tissue Gluose Uptake During Clamp

```{r bat}
plot <- barplot(summary.data$BAT_mean, 
                names.arg=summary.data$Genotype,  
                ylab="2-Deoxyglucose Uptake (ug/g/min)",
                main="Brown Adipose",
                ylim=c(0,300),
                las=1)
superpose.eb(plot, summary.data$BAT_mean, summary.data$BAT_se)
```

##  Tissue Glucose Uptake Summary

```{r tissue-glucose-uptake-summary}
par(mfrow=c(1,4))

plot <- barplot(summary.data$Gastroc_mean, 
                names.arg=(summary.data$Genotype),  
                ylab="2-Deoxyglucose Uptake (ug/g/min)",
                main="Gastrocnemius",
                ylim=c(0,17),
                las=1,
                col=grey.colors(2))
superpose.eb(plot, summary.data$Gastroc_mean, summary.data$Gastroc_se)

plot <- barplot(summary.data$BAT_mean, 
                names.arg=summary.data$Genotype,  
                ylab="2-Deoxyglucose Uptake (ug/g/min)",
                main="Brown Adipose",
                ylim=c(0,300),
                las=1,
                col=grey.colors(2))
superpose.eb(plot, summary.data$BAT_mean, summary.data$BAT_se)

plot <- barplot(summary.data$S.fat_mean, 
                names.arg=summary.data$Genotype,  
                ylab="2-Deoxyglucose Uptake (ug/g/min)",
                main="Subcutaneous\n Adipose",
                ylim=c(0,26),
                las=1,
                col=grey.colors(2))
superpose.eb(plot, summary.data$S.fat_mean, summary.data$S.fat_se)

plot <- barplot(summary.data$V.fat_mean, 
                names.arg=summary.data$Genotype,  
                ylab="2-Deoxyglucose Uptake (ug/g/min)",
                main="Visceral Adipose",
                ylim=c(0,9),
                las=1,
                col=grey.colors(2))
superpose.eb(plot, summary.data$V.fat_mean, summary.data$V.fat_se)
```

# Session Information

```{r session-information, echo=T}
sessionInfo()
```