---
title: "Weight Analysis of Dexamethasone Treated C57BL/6J Mice on a High Protein Diet"
author: "Innocence Harvey, JeAnna Redd and Dave Bridges"
date: "February 15, 2015"
output:
  html_document:
    keep_md: yes
  pdf_document:
    keep_tex: yes
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.path='figures/',
               echo=FALSE, warning=FALSE, message=FALSE,dev=c('png','pdf'))
#options(scipen = 3, digits = 2)

color.set <- c('#f77f00','#006a4d')
```

# Data Entry
This was from combined weights over several measurements of C57BL/6J Mice on treated placed on a high protein or control diet.  Some animals may appear multiple times in this analysis.  This data includes fed weights from two cohorts

```{r data-entry}
raw_data_file <- "Body Weights and Composition.csv"

data <- read.csv(raw_data_file)
data$Treatment <- relevel(data$Treatment, ref='Water')
data$Diet <- relevel(data$Diet, ref='Control Diet')
data$age.from.start <- data$age - 70
#remove ITT
data <- droplevels(subset(data, experiment.feeding_state=='fed'&assay.assay!='Grip Strength (4 Paw)'))
data$Weight <- as.numeric(as.character(data$values))/1000

library(dplyr)
library(tidyr)
weight.data <-
  subset(data, age<150&animal.Gender=="M"&Treatment!="Dexamethasone") %>%
  select(assay.assay,animal.id,age,Treatment,Diet,Weight,animal.Gender) %>%
    distinct(animal.id,assay.assay,age) %>%
  spread(assay.assay, Weight) %>%
  rename(Total.Fat.Mass = `Total Fat Mass`,
         Body.Weight = `Body Weight`,
         Lean.Mass = `Lean Mass`) %>%
  mutate(Percent.Fat.Mass = Total.Fat.Mass/Body.Weight*100)

se <- function(x) sd(x, na.rm=T)/sqrt(length(x))
combined.summary <-
  subset(weight.data, age>=70) %>%
  mutate(Weeks = round((age-70)/7)) %>%
  group_by(Weeks, Diet)%>%
  summarise(Body.Weight.mean = mean(Body.Weight, na.rm=T),
            Body.Weight.se = se(Body.Weight),
            Total.Fat.Mass.mean = mean(Total.Fat.Mass, na.rm=T),
            Total.Fat.Mass.se = se(Total.Fat.Mass),
            Lean.Mass.mean = mean(Lean.Mass, na.rm=T),
            Lean.Mass.se = se(Lean.Mass),
            Percent.Fat.Mass.mean = mean(Percent.Fat.Mass, na.rm=T),
            Percent.Fat.Mass.se = se(Percent.Fat.Mass),
            n = length(Lean.Mass)) 

superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
  arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
  length = length, ...)
```

Data was downloaded from MouseDB then aand the data is saved as `r raw_data_file`.  These data are located in `r getwd()` and was most recently updated on `r date()`.

## Body Weights

```{r body-weights}
with(subset(weight.data, Diet == levels(Diet)[1]), plot(age, Body.Weight,
                       ylab="Body Weight (g)",
                       xlab='Age (days)',
                        las=1, pch=19))
with(subset(weight.data, Diet == levels(Diet)[2]),
     points(age, Body.Weight, col=palette()[2], pch=19))

lines(with(subset(weight.data, Diet == levels(Diet)[1]), 
           loess.smooth(age,Body.Weight)), col=palette()[1])
lines(with(subset(weight.data, Diet == levels(Diet)[2]), 
           loess.smooth(age,Body.Weight)), col=palette()[2])

legend("topleft", levels(combined.summary$Diet), col=palette()[1:2], lty=1, bty='n')

yaxes <- with(combined.summary, 
              c(min(Body.Weight.mean, na.rm=T)-max(Body.Weight.se, na.rm=T), 
               c(max(Body.Weight.mean, na.rm=T)+max(Body.Weight.se, na.rm=T))))
plot <- with(subset(combined.summary, Diet==levels(Diet)[1]), 
             plot(Weeks, Body.Weight.mean, type="l", ylim=yaxes,
                  las=1, ylab="Body Weight (g)", col=palette()[1]))
with(subset(combined.summary, Diet==levels(Diet)[2]), 
             lines(Weeks, Body.Weight.mean, col=palette()[2]))

with(subset(combined.summary, Diet==levels(Diet)[1]&Weeks<=12), 
             superpose.eb(Weeks, Body.Weight.mean, Body.Weight.se, col=palette()[1]))
with(subset(combined.summary, Diet==levels(Diet)[2]&Weeks<=12), 
             superpose.eb(Weeks, Body.Weight.mean, Body.Weight.se, col=palette()[2]))
legend("topleft", levels(combined.summary$Diet), col=palette()[1:2], lty=1, bty='n')

```

```{r diagnostics-body-weight}
#statistics
library(lme4)
weight.cd.hpd.lme <- lmer(Body.Weight~age + Diet:age + Diet + (1|animal.id) + (age-1|animal.id), data=weight.data, REML=F)
weight.cd.hpd.lme.null <- lmer(Body.Weight~age + Diet + (1|animal.id) + (age-1|animal.id), data=weight.data, REML=F)
#using 
library(influence.ME)
infl <- influence(weight.cd.hpd.lme, group='animal.id')
par(mfrow=c(1,3))
plot(density(residuals(weight.cd.hpd.lme)), main="Effects of High Protein Diet", las=1)
plot(fitted(weight.cd.hpd.lme),residuals(weight.cd.hpd.lme),  
     ylab="Residuals", xlab="Fitted Values (g)", pch=19, las=1)
abline(h=0, lty=2, col='red')
barplot(cooks.distance(infl), las=2, names.arg=rownames(cooks.distance(infl)), 
         beside=T, ylab="Cook's Distance", col='grey')
```

Based on linear fixed effects models, allowing for uncorrelated intercepts and slopes for the animals, within the water group, the High Protein Diet animals gained `r fixef(weight.cd.hpd.lme)['age:DietHigh Protein Diet']*7`g  weight per week, or `r fixef(weight.cd.hpd.lme)['age:DietHigh Protein Diet']*7*12`g over the experiment.  This is a `r fixef(weight.cd.hpd.lme)['age:DietHigh Protein Diet']/fixef(weight.cd.hpd.lme)['age']*100`% change in body weight gain.  This effect was significant with a p-value of `r anova(weight.cd.hpd.lme, weight.cd.hpd.lme.null)$"Pr(>Chisq)"[2]`.  The residuals of this model **did not meet** the presumption of normality via a Shapiro-Wilk test (p=`r shapiro.test(residuals(weight.cd.hpd.lme))$p.value`).

# Lean Mass

```{r lean-mass-scatterplot}
with(subset(weight.data, Diet == levels(Diet)[1]), plot(age, Lean.Mass,
                       ylab="Total Lean Mass (g)",
                       xlab='Age (days)',
                        las=1, pch=19))
with(subset(weight.data, Diet == levels(Diet)[2]),
     points(age, Lean.Mass, col=palette()[2], pch=19))
se <- function(x) sd(x)/sqrt(length(x))

lines(with(subset(weight.data, Diet == levels(Diet)[1]), 
           loess.smooth(age,Lean.Mass)), col=palette()[1])
lines(with(subset(weight.data, Diet == levels(Diet)[2]), 
           loess.smooth(age,Lean.Mass)), col=palette()[2])

legend("topleft", levels(combined.summary$Diet), col=palette()[1:2], lty=1, bty='n')

yaxes <- with(combined.summary, 
              c(min(Lean.Mass.mean, na.rm=T)-max(Lean.Mass.se, na.rm=T), 
               c(max(Lean.Mass.mean, na.rm=T)+max(Lean.Mass.se, na.rm=T))))
plot <- with(subset(combined.summary, Diet==levels(Diet)[1]), 
             plot(Weeks, Lean.Mass.mean, type="l", ylim=yaxes,
                  las=1, ylab="Total Lean Mass (g)", col=palette()[1]))
with(subset(combined.summary, Diet==levels(Diet)[2]), 
             lines(Weeks, Lean.Mass.mean, col=palette()[2]))

with(subset(combined.summary, Diet==levels(Diet)[1]&Weeks<=12), 
             superpose.eb(Weeks, Lean.Mass.mean, Lean.Mass.se, col=palette()[1]))
with(subset(combined.summary, Diet==levels(Diet)[2]&Weeks<=12), 
             superpose.eb(Weeks, Lean.Mass.mean, Lean.Mass.se, col=palette()[2]))
legend("topleft", levels(combined.summary$Diet), col=palette()[1:2], lty=1, bty='n')

#statistics
library(lme4)
lean.lme <- lmer(Lean.Mass~age*Treatment*Diet + (1|animal.id) + (age-1|animal.id), data=weight.data, REML=F)
lean.lme.null <- lmer(Lean.Mass~age*Treatment + (1|animal.id) + (age-1|animal.id), data=weight.data, REML=F)
```

Based on a mixed linear model with a random slope and intercept for each animal, there was a significant decrease in the rate of lean mass gain with respect to time (Chisq = `r anova(lean.lme,lean.lme.null)$Chisq[2]`, p=`r anova(lean.lme,lean.lme.null)$"Pr(>Chisq)"[2]`).  This as an average difference of `r coefficients(summary(lean.lme))['age:DietHigh Protein Diet','Estimate']*7` +/- `r coefficients(summary(lean.lme))['age:DietHigh Protein Diet','Std. Error']*7` grams fat per week, or a total of `r coefficients(summary(lean.lme))['age:DietHigh Protein Diet','Estimate']*180` +/- `r coefficients(summary(lean.lme))['age:DietHigh Protein Diet','Std. Error']*180` over the course of the study.  This is a `r -coefficients(summary(lean.lme))['age:DietHigh Protein Diet','Estimate']/coefficients(summary(lean.lme))['age','Estimate']*100`% reduction.

# Fat Mass

```{r fat-mass-scatterplot}
with(subset(weight.data, Diet == levels(Diet)[1]), plot(age, Total.Fat.Mass,
                       ylab="Total Fat Mass (g)",
                       xlab='Age (days)',
                        las=1, pch=19))
with(subset(weight.data, Diet == levels(Diet)[2]),
     points(age, Total.Fat.Mass, col=palette()[2], pch=19))

lines(with(subset(weight.data, Diet == levels(Diet)[1]), 
           loess.smooth(age,Total.Fat.Mass)), col=palette()[1])
lines(with(subset(weight.data, Diet == levels(Diet)[2]), 
           loess.smooth(age,Total.Fat.Mass)), col=palette()[2])

legend("topleft", levels(combined.summary$Diet), col=palette()[1:2], lty=1, bty='n')

yaxes <- with(combined.summary, 
              c(min(Total.Fat.Mass.mean, na.rm=T)-max(Total.Fat.Mass.se, na.rm=T), 
               c(max(Total.Fat.Mass.mean, na.rm=T)+max(Total.Fat.Mass.se, na.rm=T))))
plot <- with(subset(combined.summary, Diet==levels(Diet)[1]), 
             plot(Weeks, Total.Fat.Mass.mean, type="l", ylim=yaxes,
                  las=1, ylab="Total Fat Mass (g)", col=palette()[1]))
with(subset(combined.summary, Diet==levels(Diet)[2]), 
             lines(Weeks, Total.Fat.Mass.mean, col=palette()[2]))

with(subset(combined.summary, Diet==levels(Diet)[1]&Weeks<=12), 
             superpose.eb(Weeks, Total.Fat.Mass.mean, Total.Fat.Mass.se, col=palette()[1]))
with(subset(combined.summary, Diet==levels(Diet)[2]&Weeks<=12), 
             superpose.eb(Weeks, Total.Fat.Mass.mean, Total.Fat.Mass.se, col=palette()[2]))
legend("topleft", levels(combined.summary$Diet), col=palette()[1:2], lty=1, bty='n')

#statistics
fat.lme <- lmer(Total.Fat.Mass~age*Diet + (1|animal.id) + (age-1|animal.id), data=weight.data, REML=F)
fat.lme.null <- lmer(Total.Fat.Mass~age + (1|animal.id) + (age-1|animal.id), data=weight.data, REML=F)
```


```{r hpd-fat-mass-scatterplot-uthsc}
plot <- with(subset(combined.summary, Diet=="Control Diet"), 
             plot(Weeks, Total.Fat.Mass.mean, type="l", ylim=yaxes,
                  cex.names=1.5,cex.lab=1.5,
                  las=1, ylab="Total Fat Mass (g)", col=color.set[1]))
with(subset(combined.summary, Diet=="High Protein Diet"), 
             lines(Weeks, Total.Fat.Mass.mean, col=color.set[2]))

with(subset(combined.summary, Diet=="Control Diet"), 
             superpose.eb(Weeks, Total.Fat.Mass.mean, Total.Fat.Mass.se, col=color.set[1]))
with(subset(combined.summary, Diet=="High Protein Diet"), 
             superpose.eb(Weeks, Total.Fat.Mass.mean, Total.Fat.Mass.se, col=color.set[2]))
legend("topleft", levels(combined.summary$Diet), col=color.set, lty=1, bty='n', cex=1.5)
```

Based on a mixed linear model with a random slope and intercept for each animal, there was a significant decrease in the rate of fat mass gain with respect to time (Chisq = `r anova(fat.lme,fat.lme.null)$Chisq[2]`, p=`r anova(fat.lme,fat.lme.null)$"Pr(>Chisq)"[2]`).  This as an average difference of `r coefficients(summary(fat.lme))['age:DietHigh Protein Diet','Estimate']*7` +/- `r coefficients(summary(fat.lme))['age:DietHigh Protein Diet','Std. Error']*7` grams fat per week, or a total of `r coefficients(summary(fat.lme))['age:DietHigh Protein Diet','Estimate']*180` +/- `r coefficients(summary(fat.lme))['age:DietHigh Protein Diet','Std. Error']*180` over the course of the study.  This is a `r -coefficients(summary(fat.lme))['age:DietHigh Protein Diet','Estimate']/coefficients(summary(fat.lme))['age','Estimate']*100`% reduction.


## Percent Fat Mass

```{r pct-fat-mass-scatterplot}
with(subset(weight.data, Diet == levels(Diet)[1]), plot(age, Percent.Fat.Mass,
                       ylab="Percent Fat Mass",
                       xlab='Age (days)',
                        las=1, pch=19))
with(subset(weight.data, Diet == levels(Diet)[2]),
     points(age, Percent.Fat.Mass, col=palette()[2], pch=19))

lines(with(subset(weight.data, Diet == levels(Diet)[1]), 
           loess.smooth(age,Percent.Fat.Mass)), col=palette()[1])
lines(with(subset(weight.data, Diet == levels(Diet)[2]), 
           loess.smooth(age,Percent.Fat.Mass)), col=palette()[2])

legend("topleft", levels(combined.summary$Diet), col=palette()[1:2], lty=1, bty='n')

yaxes <- with(combined.summary, 
              c(min(Percent.Fat.Mass.mean, na.rm=T)-max(Percent.Fat.Mass.se, na.rm=T), 
               c(max(Percent.Fat.Mass.mean, na.rm=T)+max(Percent.Fat.Mass.se, na.rm=T))))
plot <- with(subset(combined.summary, Diet==levels(Diet)[1]), 
             plot(Weeks, Percent.Fat.Mass.mean, type="l", ylim=yaxes,
                  las=1, ylab="Percent Fat Mass", col=palette()[1]))
with(subset(combined.summary, Diet==levels(Diet)[2]), 
             lines(Weeks, Percent.Fat.Mass.mean, col=palette()[2]))

with(subset(combined.summary, Diet==levels(Diet)[1]&Weeks<=12), 
             superpose.eb(Weeks, Percent.Fat.Mass.mean, Percent.Fat.Mass.se, col=palette()[1]))
with(subset(combined.summary, Diet==levels(Diet)[2]&Weeks<=12), 
             superpose.eb(Weeks, Percent.Fat.Mass.mean, Percent.Fat.Mass.se, col=palette()[2]))
legend("topleft", levels(combined.summary$Diet), col=palette()[1:2], lty=1, bty='n')

#statistics
pct.fat.lme <- lmer(Percent.Fat.Mass~age*Diet + (1|animal.id) + (age-1|animal.id), data=weight.data, REML=F)
pct.fat.lme.null <- lmer(Percent.Fat.Mass~age + (1|animal.id) + (age-1|animal.id), data=weight.data, REML=F)
```

Based on a mixed linear model with a random slope and intercept for each animal, there was a significant decrease in the rate of percent fat mass with respect to time (Chisq = `r anova(pct.fat.lme,pct.fat.lme.null)$Chisq[2]`, p=`r anova(pct.fat.lme,pct.fat.lme.null)$"Pr(>Chisq)"[2]`).  This as an average difference of `r coefficients(summary(pct.fat.lme))['age:DietHigh Protein Diet','Estimate']*7` +/- `r coefficients(summary(pct.fat.lme))['age:DietHigh Protein Diet','Std. Error']*7` percent fat per week, or a total of `r coefficients(summary(pct.fat.lme))['age:DietHigh Protein Diet','Estimate']*180` +/- `r coefficients(summary(pct.fat.lme))['age:DietHigh Protein Diet','Std. Error']*180` over the course of the study.  This is a `r -coefficients(summary(pct.fat.lme))['age:DietHigh Protein Diet','Estimate']/coefficients(summary(pct.fat.lme))['age','Estimate']*100`% reduction.


# Session Information
```{r sessionInfo}
sessionInfo()
```

