---
title: "Tissue Weight Analysis of HPD Fed Animals"
author: "Innocence Harvey and Dave Bridges"
date: "August 10, 2015"
output:
  html_document:
    keep_md: yes
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.path='figures/',
               echo=FALSE, warning=FALSE, message=FALSE,dev=c('png','pdf'))
options(scipen=3,digits = 3)

color.set <- c('#f77f00','#006a4d')
```

```{r data-entry}
filename <- 'HPD fat pad weights.csv'
data <- read.csv(filename)
data$Diet <- relevel(data$Diet, ref="Control Diet")
```

This data uses the input file `r filename`.  This script was most recently run on `r date()`.

# Adipose Tissue

The iWAT and eWAT values are the sum of the left and right fat pads for each mouse.

```{r hpd-wat}
data$eWAT <- data$Right.EWAT + data$Left.EWAT
data$iWAT <- data$Right.IWAT + data$Left.IWAT

se <- function(x) sd(x, na.rm=T)/sqrt(length(x))
library(dplyr)
iwat.summary <- 
  data %>%
  select(Diet,Body.Weight,iWAT) %>%
  group_by(Diet) %>%
  summarise(
    n = length(iWAT),
    Average = mean(iWAT, na.rm=T),
    SD = sd(iWAT, na.rm=T),
    SE = se(iWAT),
    Shapiro = shapiro.test(iWAT)$p.value)

ewat.summary <- 
  data %>%
  select(Diet,Body.Weight,eWAT) %>%
  group_by(Diet) %>%
  summarise(
    n = length(eWAT),
    Average = mean(eWAT, na.rm=T),
    SD = sd(eWAT, na.rm=T),
    SE = se(eWAT),
    Shapiro = shapiro.test(eWAT)$p.value)
library(car)
```

We tested whether normality can be assumed for eWAT and iWAT by a Shapiro-Wilk test.  The p-values for this were all greater than `r min(ewat.summary$Shapiro, iwat.summary$Shapiro)`.  For summary statistics see below.

```{r adipose-summary-tables}
kable(iwat.summary, caption="Inguinal Fat Pads (mg)")
kable(ewat.summary, caption="Epididymal Fat Pads (mg)")
```

```{r hpd-adipose-weights}
superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
  arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
  length = length, ...)

ymax <- max(ewat.summary$Average + ewat.summary$SE)
plot <- barplot(cbind(iWAT=iwat.summary$Average, eWAT=ewat.summary$Average),
                beside=T, las=1, ylab="Fat Pad Mass (mg)",
                ylim=c(0,ymax))
superpose.eb(plot,
             cbind(iWAT=iwat.summary$Average, eWAT=ewat.summary$Average),
             cbind(iWAT=iwat.summary$SE, eWAT=ewat.summary$SE))
legend('topleft', levels(data$Diet), fill=grey.colors(2), bty="n")
```

```{r hpd-adipose-weights-uthsc}
ymax <- max(ewat.summary$Average + ewat.summary$SE)
plot <- barplot(cbind(iWAT=iwat.summary$Average, eWAT=ewat.summary$Average),
                beside=T, las=1, ylab="Fat Pad Mass (mg)",
                cex.lab=1.5, cex.names=1.5, col=color.set,
                ylim=c(0,ymax))
superpose.eb(plot,
             cbind(iWAT=iwat.summary$Average, eWAT=ewat.summary$Average),
             cbind(iWAT=iwat.summary$SE, eWAT=ewat.summary$SE))
legend('topleft', levels(data$Diet), fill=color.set, bty="n",cex=1.5)
```

We next tested whether equal variance can be assumed via Levene's test.  For iWAT the p-value was `r leveneTest(iWAT~Diet, data=data)$"Pr(>F)"[1]` and for eWAT the p-value was `r leveneTest(eWAT~Diet, data=data)$"Pr(>F)"[1]`.  Therefore in both cases, equal variance can not be assumed and Welch's *t* tests were performed.  These p-values were **`r t.test(iWAT~Diet, var.equal=F, data=data)$p.value`** for iWAT and **`r t.test(eWAT~Diet, var.equal=F, data=data)$p.value`** for eWAT.

# Muscle Tissues

The TS and Quadriceps values are the sum of the left and right fat pads for each mouse.

```{r hpd-muscle}
data$Quad <- data$Right.Quad + data$Left.Quad
data$TS <- data$TS.Right + data$TS.Left

ts.summary <- 
  data %>%
  select(Diet,Body.Weight,TS) %>%
  group_by(Diet) %>%
  summarise(
    n = length(TS),
    Average = mean(TS, na.rm=T),
    SD = sd(TS, na.rm=T),
    SE = se(TS),
    Shapiro = shapiro.test(TS)$p.value)

quad.summary <- 
  data %>%
  select(Diet,Body.Weight,Quad) %>%
  group_by(Diet) %>%
  summarise(
    n = length(Quad),
    Average = mean(Quad, na.rm=T),
    SD = sd(Quad, na.rm=T),
    SE = se(Quad),
    Shapiro = shapiro.test(Quad)$p.value)

heart.summary <- 
  data %>%
  select(Diet,Body.Weight,Heart) %>%
  group_by(Diet) %>%
  summarise(
    n = length(Heart),
    Average = mean(Heart, na.rm=T),
    SD = sd(Heart, na.rm=T),
    SE = se(Heart),
    Shapiro = shapiro.test(Heart)$p.value)

```

We tested whether normality can be assumed for TS, Quadriceps and Heart by Shapiro-Wilk tests.  The p-values for this were all greater than `r min(ts.summary$Shapiro, quad.summary$Shapiro, heart.summary$Shapiro)`.  For summary statistics see below.

```{r muscle-summary-tables}
kable(ts.summary, caption="Triceps Surae Weights (mg)")
kable(quad.summary, caption="Quadriceps Weights (mg)")
kable(heart.summary, caption="Heart Weights (mg)")
```

```{r hpd-muscle-weights}
ymax <- max(quad.summary$Average + quad.summary$SE)
plot <- barplot(cbind(Quadriceps=quad.summary$Average,
                      `Triceps surae`=ts.summary$Average,
                      Heart = heart.summary$Average),
                beside=T, las=1, ylab=" Muscle Weight (mg)",
                ylim=c(0,ymax))
superpose.eb(plot,
             cbind(Quadriceps=quad.summary$Average,
                      `Triceps surae`=ts.summary$Average,
                      Heart = heart.summary$Average),
             cbind(Quadriceps=quad.summary$SE,
                      `Triceps surae`=ts.summary$SE,
                      Heart = heart.summary$SE))
legend('topright', levels(data$Diet), fill=grey.colors(2), bty="n")
```

```{r hpd-muscle-weights-uthsc}
ymax <- max(quad.summary$Average + quad.summary$SE)
plot <- barplot(cbind(Quadriceps=quad.summary$Average,
                      `Triceps surae`=ts.summary$Average,
                      Heart = heart.summary$Average),
                beside=T, las=1, ylab=" Muscle Weight (mg)",
                cex.names=1.5, cex.lab=1.5,col=color.set,
                ylim=c(0,ymax))
superpose.eb(plot,
             cbind(Quadriceps=quad.summary$Average,
                      `Triceps surae`=ts.summary$Average,
                      Heart = heart.summary$Average),
             cbind(Quadriceps=quad.summary$SE,
                      `Triceps surae`=ts.summary$SE,
                      Heart = heart.summary$SE))
legend('topright', levels(data$Diet), fill=color.set, bty="n", cex=1.5)
```


We next tested whether equal variance can be assumed via Levene's test.  For quadriceps the p-value was `r leveneTest(Quad~Diet, data=data)$"Pr(>F)"[1]`,  for TS the p-value was `r leveneTest(TS~Diet, data=data)$"Pr(>F)"[1]` and for heart the p-value was `r leveneTest(Heart~Diet, data=data)$"Pr(>F)"[1]`.  Therefore in all cases, equal variance can be assumed and Student's *t* tests were performed.  These p-values were **`r t.test(Quad~Diet, var.equal=T, data=data)$p.value`** for quadriceps, **`r t.test(TS~Diet, var.equal=T, data=data)$p.value`** for TS, and **`r t.test(Heart~Diet, var.equal=T, data=data)$p.value`** for heart.


# Session Information

```{r session-info, echo=FALSE, message=FALSE}
sessionInfo()
```