---
title: "Food Intake Analysis of HPD Fed Animals"
author: "Innocence Harvey and Dave Bridges"
date: "August 10, 2015"
output:
  html_document:
    keep_md: yes
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.path='figures/',
               echo=FALSE, warning=FALSE, message=FALSE,dev=c('png','pdf'))
options(scipen=3,digits = 3)

color.set <- c('#f77f00','#006a4d')
```

```{r data-entry}
filename <- 'HPD Food Intake Data.csv'
data <- read.csv(filename)
data$Food.Intake <- data$Food.Intake..g.
#remove dex measurements
library(dplyr)
food.data <-
  subset(data, Treatment %in% c("CD","HPD")) %>%
  select(Week, Food, Treatment) %>%
  mutate(Week = floor(Week))
food.data <- droplevels(food.data)
food.data$Cage <- as.factor(rep(c(1,2,3,4),12))

#caloric intake is 3.8 kcal/g for both groups
calories <- 3.8
se <- function(x) sd(x)/sqrt(length(x))
food.summary <- 
  food.data %>%
  group_by(Week,Treatment) %>%
  summarize(Average = mean(Food),
            SD = sd(Food),
            SE = se(Food))

kcal.summary <- data.frame(
  Week = food.summary$Week,
  Treatment = food.summary$Treatment,
  Average = food.summary$Average*calories,
  SD = food.summary$SD*calories,
  SE = food.summary$SE*calories
)
```

This data uses the input file `r filename`.  This script was most recently run on `r date()`.


```{r hpd-food-intake-grams, fig.cap="Food Intake Per Week"}
ymax <- max(food.summary$Average + food.summary$SE, na.rm=T)
plot <- with(subset(food.summary, Treatment=="CD"), plot(Week,Average, las=1,
     ylim=c(0,ymax), type="l",ylab="Food Intake (g)"))
with(subset(food.summary, Treatment=="HPD"), lines(Week,Average, col="red"))
legend("bottomleft", levels(food.data$Treatment), bty="n", lty=1, col=palette()[1:2])

superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
  arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
  length = length, ...)

with(subset(food.summary, Treatment=="CD"), superpose.eb(Week,Average,SE, col=palette()[1]))
with(subset(food.summary, Treatment=="HPD"), superpose.eb(Week,Average,SE, col=palette()[2]))
```

```{r hpd-food-intake-kcal, fig.cap="Food Intake Per Week"}
ymax <- max(kcal.summary$Average + kcal.summary$SE, na.rm=T)
plot <- with(subset(kcal.summary, Treatment=="CD"), plot(Week,Average, las=1,
     ylim=c(0,ymax), type="l",ylab="Food Intake (kCal)"))
with(subset(kcal.summary, Treatment=="HPD"), lines(Week,Average, col="red"))
legend("bottomleft", levels(kcal.summary$Treatment), bty="n", lty=1, col=palette()[1:2])

with(subset(kcal.summary, Treatment=="CD"), superpose.eb(Week,Average,SE, col=palette()[1]))
with(subset(kcal.summary, Treatment=="HPD"), superpose.eb(Week,Average,SE, col=palette()[2]))
```

```{r hpd-food-intake-kcal-uthsc, fig.cap="Food Intake Per Week"}
par(mar=c(5, 7, 4, 2))
ymax <- max(kcal.summary$Average + kcal.summary$SE, na.rm=T)
plot <- with(subset(kcal.summary, Treatment=="CD"), 
             plot(Week,Average, las=1,cex.names=1.5,
                  cex.lab=1.5,
                  ylim=c(0,ymax), type="l",
                  ylab="Food Intake (kCal)", col=color.set[1]))
with(subset(kcal.summary, Treatment=="HPD"), lines(Week,Average, col=color.set[2]))
legend("bottomleft", levels(kcal.summary$Treatment), bty="n", lty=1, col=color.set, cex=1.5)

with(subset(kcal.summary, Treatment=="CD"), superpose.eb(Week,Average,SE, col=color.set[1]))
with(subset(kcal.summary, Treatment=="HPD"), superpose.eb(Week,Average,SE, col=color.set[2]))
```

## Statistics

```{r food-intake-statistics}
library(lme4)
food.lme <- lmer(Food ~ Treatment + (1|Cage), REML=F, data=food.data)
food.lme.null <- lmer(Food ~ 1 + (1|Cage), REML=F, data=food.data)
```

To test the effects of food intake we used a mixed linear model in which the week did not matter and either the Treatment covariate was included or not.  The cage was the random effect.   From this model the Chi-squared value was `r  anova(food.lme, food.lme.null)$Chisq[2]` for a p-value of `r anova(food.lme, food.lme.null)$"Pr(>Chisq)"[2]`.  The residuals of this model were normally distributed (Shapiro-Wilk test p-value = `r shapiro.test(residuals(food.lme))$p.value`).

# Cumulative Food Intake

```{r hpd-food-intake-kcal-cumulative, fig.cap="Cumulative Food Intake"}
cum.food.data <- 
  food.data[complete.cases(food.data),] %>%
  group_by(Cage,Treatment,Week) %>%
  summarize(Food.Intake = sum(Food)) %>%
  mutate(Cumulative = cumsum(Food.Intake))

cumulative.summary <- 
  cum.food.data %>%
  group_by(Week,Treatment) %>%
  summarize(Average = mean(Cumulative),
            SD = sd(Cumulative),
            SE = se(Cumulative))

ymax <- max(cumulative.summary$Average + cumulative.summary$SE, na.rm=T)
plot <- with(subset(cumulative.summary, Treatment=="CD"), 
             plot(Week,Average, las=1,
     ylim=c(0,ymax), type="l",ylab="Food Intake (kCal)"))
with(subset(cumulative.summary, Treatment=="HPD"), lines(Week,Average, col="red"))
legend("topleft", levels(cumulative.summary$Treatment), bty="n", lty=1, col=palette()[1:2])

with(subset(cumulative.summary, Treatment=="CD"), superpose.eb(Week,Average,SE, col=palette()[1]))
with(subset(cumulative.summary, Treatment=="HPD"), superpose.eb(Week,Average,SE, col=palette()[2]))
```

# Session Information

```{r session-info, echo=FALSE, message=FALSE}
sessionInfo()
```