---
title: "mTORC1 Signaling in HPD Quadriceps"
author: "Dave Bridges"
date: "June 26, 2015"
output:
  html_document:
    keep_md: yes
---


```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.path='figures/',
               echo=FALSE, warning=FALSE, message=FALSE,dev=c('png','pdf'))

options(scipen=8, digits=3)
```

# S6K Phosphorylation

```{r s6k-analysis,  fig.cap="Normalized mTORC1 Activity in Quadriceps"}
datafile <- 'HPD Quadriceps Blot Quantification.csv'
library(readr)
data <- read_csv(file=datafile)

s6k.data <- subset(data, `Image Name`=='S6K')
#select band 1 for phospho
p.s6k.data <- s6k.data[s6k.data$Channel==700,]
p.s6k.data$Protein <- rep("pS6K", dim(p.s6k.data)[1])

p.s6k.data$Treatment <- c(rep("Fasted",8), rep("Fed",2))
p.s6k.data$Diet <- c(rep("CD", 4), rep("HPD",4), rep("CD",2))
p.s6k.data$Lane <- as.factor(seq(1,10))

t.s6k.data <- s6k.data[s6k.data$Channel==800,]
t.s6k.data$Protein <- rep("S6K", dim(t.s6k.data)[1])
t.s6k.data$Diet <- c(rep("CD", 4), rep("HPD",4), rep("CD",2))
t.s6k.data$Treatment <- c(rep("Fasted",8), rep("Fed",2))
t.s6k.data$Lane <- as.factor(seq(1,10))

library(dplyr)
library(tidyr)
s6k.summary.lane <- 
  rbind(p.s6k.data,t.s6k.data) %>%
  mutate(Intensity = Signal) %>%
  spread(Protein, Intensity) %>%
  group_by(Treatment,Diet,Lane) %>%
  summarize(pS6K = mean(pS6K, na.rm=T),
            S6K = mean(S6K, na.rm=T)) %>%
  mutate(Ratio=pS6K/S6K)
s6k.summary.lane$Norm.Ratio <- s6k.summary.lane$Ratio/mean(s6k.summary.lane$Ratio[1:4])


#remove outlier lane
#s6k.summary.lane <- subset(s6k.summary.lane,!( `Lane Name` %in% c("Lane 2","Lane 8")))
kable(s6k.summary.lane, caption="Lane Summary for S6K", digits=2)

s6k.summary <-
  s6k.summary.lane %>%
  group_by(Treatment, Diet) %>%
  summarize(Average = mean(Norm.Ratio),
            SE = sd(Norm.Ratio)/sqrt(length(Norm.Ratio)),
            SD = sd(Norm.Ratio),
            n = length(Norm.Ratio))

kable(s6k.summary, caption="Summarized Normalized Ratios for S6K", digits=2)
s6k.plot <-
  s6k.summary %>%
  select(Diet,Treatment,Average) %>%
  spread(Treatment, Average)
s6k.plot.se <-
  s6k.summary %>%
  select(Diet,Treatment,SE) %>%
  spread(Treatment, SE)

ymax <- max(s6k.summary$Average + s6k.summary$SE)
plot <- barplot(as.matrix(s6k.plot[2:3]), beside=T, las=1,
        ylab="pS6K/S6K", ylim=c(0,ymax))

superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
  arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
  length = length, ...)

superpose.eb(plot,
             as.matrix(s6k.plot[2:3]),
             as.matrix(s6k.plot.se[2:3]))
```

## S6K Statistics

```{r s6k-statistics}
s6k.aov <- aov(Norm.Ratio~Treatment+Diet, data=s6k.summary.lane)
#kable(summary(s6k.aov)[[1]], caption="Two Way ANOVA for Effects of Diet and Treatment on Normalized pS6K Ratio", digits=3)
library(car)
```

Ignoring the fed mice, within the fasted mice, the data was normally distributed (Shapiro-Wilk test `r with(s6k.summary.lane, shapiro.test(s6k.summary.lane[Diet=="CD"&Treatment=="Fasted",]$Norm.Ratio))$p.value` for CD and `r with(s6k.summary.lane, shapiro.test(s6k.summary.lane[Diet=="HPD"&Treatment=="Fasted",]$Norm.Ratio))$p.value` for HPD), and equal variance could be assumed (Levene's test `r leveneTest(Norm.Ratio~Diet, subset(s6k.summary.lane, Treatment=="Fasted"))$"Pr(>F)"[1]`).  Based on this a Student's *t* test had a p-value of `r t.test(Norm.Ratio~Diet, subset(s6k.summary.lane, Treatment=="Fasted"), var.equal=T)$p.value`.

# Akt Phosphorylation

```{r akt-analysis,  fig.cap="Normalized Akt Activity in Quadriceps"}

Akt.data <- subset(data, `Image Name`=='Akt')
#select band 1 for phospho
p.Akt.data <- Akt.data[Akt.data$Channel==700,]
p.Akt.data$Protein <- rep("pAkt", dim(p.Akt.data)[1])

p.Akt.data$Treatment <- c(rep("Fasted",8), rep("Fed",2))
p.Akt.data$Diet <- c(rep("CD", 4), rep("HPD",4), rep("CD",2))
p.Akt.data$Lane <- as.factor(seq(1,10))

t.Akt.data <- Akt.data[Akt.data$Channel==800,]
t.Akt.data$Protein <- rep("Akt", dim(t.Akt.data)[1])
t.Akt.data$Diet <- c(rep("CD", 4), rep("HPD",4), rep("CD",2))
t.Akt.data$Treatment <- c(rep("Fasted",8), rep("Fed",2))
t.Akt.data$Lane <- as.factor(seq(1,10))

Akt.summary.lane <- 
  rbind(p.Akt.data,t.Akt.data) %>%
  mutate(Intensity = Signal) %>%
  spread(Protein, Intensity) %>%
  group_by(Treatment,Diet,Lane) %>%
  summarize(pAkt = mean(pAkt, na.rm=T),
            Akt = mean(Akt, na.rm=T)) %>%
  mutate(Ratio=pAkt/Akt)
Akt.summary.lane$Norm.Ratio <- Akt.summary.lane$Ratio/mean(Akt.summary.lane$Ratio[1:4])


#remove outlier lane
#Akt.summary.lane <- subset(Akt.summary.lane,!( `Lane Name` %in% c("Lane 2","Lane 8")))
kable(Akt.summary.lane, caption="Lane Summary for Akt", digits=2)

Akt.summary <-
  Akt.summary.lane %>%
  group_by(Treatment, Diet) %>%
  summarize(Average = mean(Norm.Ratio),
            SE = sd(Norm.Ratio)/sqrt(length(Norm.Ratio)),
            SD = sd(Norm.Ratio),
            n = length(Norm.Ratio))

kable(Akt.summary, caption="Summarized Normalized Ratios for Akt", digits=2)
Akt.plot <-
  Akt.summary %>%
  select(Diet,Treatment,Average) %>%
  spread(Treatment, Average)
Akt.plot.se <-
  Akt.summary %>%
  select(Diet,Treatment,SE) %>%
  spread(Treatment, SE)

ymax <- max(Akt.summary$Average + Akt.summary$SE)
plot <- barplot(as.matrix(Akt.plot[2:3]), beside=T, las=1,
        ylab="pAkt/Akt", ylim=c(0,ymax))

superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
  arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
  length = length, ...)

superpose.eb(plot,
             as.matrix(Akt.plot[2:3]),
             as.matrix(Akt.plot.se[2:3]))
```

## Akt Statistics

```{r Akt-statistics}
Akt.aov <- aov(Norm.Ratio~Treatment+Diet, data=Akt.summary.lane)
#kable(summary(Akt.aov)[[1]], caption="Two Way ANOVA for Effects of Diet and Treatment on Normalized pAkt Ratio", digits=3)
library(car)
```

Ignoring the fed mice, within the fasted mice, the data was normally distributed (Shapiro-Wilk test `r with(Akt.summary.lane, shapiro.test(Akt.summary.lane[Diet=="CD"&Treatment=="Fasted",]$Norm.Ratio))$p.value` for CD and `r with(Akt.summary.lane, shapiro.test(Akt.summary.lane[Diet=="HPD"&Treatment=="Fasted",]$Norm.Ratio))$p.value` for HPD), and equal variance could be assumed (Levene's test `r leveneTest(Norm.Ratio~Diet, subset(Akt.summary.lane, Treatment=="Fasted"))$"Pr(>F)"[1]`).  Based on this a Student's *t* test had a p-value of `r t.test(Norm.Ratio~Diet, subset(Akt.summary.lane, Treatment=="Fasted"), var.equal=T)$p.value`.

```{r tsc-analysis,  fig.cap="Normalized TSC Phosphorylation in Quadriceps"}

tsc.data <- subset(data, `Image Name`=='TSC')
#select band 1 for phospho
p.tsc.data <- tsc.data[tsc.data$Channel==800,]
p.tsc.data$Protein <- rep("ptsc", dim(p.tsc.data)[1])

p.tsc.data$Treatment <- c(rep("Fasted",8), rep("Fed",2))
p.tsc.data$Diet <- c(rep("CD", 4), rep("HPD",4), rep("CD",2))
p.tsc.data$Lane <- as.factor(seq(1,10))

t.tsc.data <- tsc.data[tsc.data$Channel==700,]
t.tsc.data$Protein <- rep("tsc", dim(t.tsc.data)[1])
t.tsc.data$Diet <- c(rep("CD", 4), rep("HPD",4), rep("CD",2))
t.tsc.data$Treatment <- c(rep("Fasted",8), rep("Fed",2))
t.tsc.data$Lane <- as.factor(seq(1,10))

tsc.summary.lane <- 
  rbind(p.tsc.data,t.tsc.data) %>%
  mutate(Intensity = Signal) %>%
  spread(Protein, Intensity) %>%
  group_by(Treatment,Diet,Lane) %>%
  summarize(ptsc = mean(ptsc, na.rm=T),
            tsc = mean(tsc, na.rm=T)) %>%
  mutate(Ratio=ptsc/tsc)
tsc.summary.lane$Norm.Ratio <- tsc.summary.lane$Ratio/mean(tsc.summary.lane$Ratio[1:4])


#remove outlier lane
#tsc.summary.lane <- subset(tsc.summary.lane,!( `Lane Name` %in% c("Lane 2","Lane 8")))
kable(tsc.summary.lane, caption="Lane Summary for TSC", digits=2)

tsc.summary <-
  tsc.summary.lane %>%
  group_by(Treatment, Diet) %>%
  summarize(Average = mean(Norm.Ratio),
            SE = sd(Norm.Ratio)/sqrt(length(Norm.Ratio)),
            SD = sd(Norm.Ratio),
            n = length(Norm.Ratio))

kable(tsc.summary, caption="Summarized Normalized Ratios for TSC2", digits=2)
tsc.plot <-
  tsc.summary %>%
  select(Diet,Treatment,Average) %>%
  spread(Treatment, Average)
tsc.plot.se <-
  tsc.summary %>%
  select(Diet,Treatment,SE) %>%
  spread(Treatment, SE)

ymax <- max(tsc.summary$Average + tsc.summary$SE)
plot <- barplot(as.matrix(tsc.plot[2:3]), beside=T, las=1,
        ylab="pTSC2/TSC2", ylim=c(0,ymax))

superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
  arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
  length = length, ...)

superpose.eb(plot,
             as.matrix(tsc.plot[2:3]),
             as.matrix(tsc.plot.se[2:3]))
```

## TSC Statistics

```{r tsc-statistics}
tsc.aov <- aov(Norm.Ratio~Treatment+Diet, data=tsc.summary.lane)
#kable(summary(tsc.aov)[[1]], caption="Two Way ANOVA for Effects of Diet and Treatment on Normalized pTSC Ratio", digits=3)
library(car)
```

Ignoring the fed mice, within the fasted mice, the data was normally distributed (Shapiro-Wilk test `r with(tsc.summary.lane, shapiro.test(tsc.summary.lane[Diet=="CD"&Treatment=="Fasted",]$Norm.Ratio))$p.value` for CD and `r with(tsc.summary.lane, shapiro.test(tsc.summary.lane[Diet=="HPD"&Treatment=="Fasted",]$Norm.Ratio))$p.value` for HPD), and equal variance could be assumed (Levene's test `r leveneTest(Norm.Ratio~Diet, subset(tsc.summary.lane, Treatment=="Fasted"))$"Pr(>F)"[1]`).  Based on this a Student's *t* test had a p-value of `r t.test(Norm.Ratio~Diet, subset(tsc.summary.lane, Treatment=="Fasted"), var.equal=T)$p.value`.

```{r s6-analysis,  fig.cap="Normalized S6 Phosphorylation in Quadriceps"}

s6.data <- subset(data, `Image Name`=='S6')
#select band 1 for phospho
p.s6.data <- s6.data[s6.data$Channel==700,]
p.s6.data$Protein <- rep("ps6", dim(p.s6.data)[1])

p.s6.data$Treatment <- c(rep("Fasted",8), rep("Fed",2))
p.s6.data$Diet <- c(rep("CD", 4), rep("HPD",4), rep("CD",2))
p.s6.data$Lane <- as.factor(seq(1,10))

t.s6.data <- s6.data[s6.data$Channel==800,]
t.s6.data$Protein <- rep("s6", dim(t.s6.data)[1])
t.s6.data$Diet <- c(rep("CD", 4), rep("HPD",4), rep("CD",2))
t.s6.data$Treatment <- c(rep("Fasted",8), rep("Fed",2))
t.s6.data$Lane <- as.factor(seq(1,10))

s6.summary.lane <- 
  rbind(p.s6.data,t.s6.data) %>%
  mutate(Intensity = Signal) %>%
  spread(Protein, Intensity) %>%
  group_by(Treatment,Diet,Lane) %>%
  summarize(ps6 = mean(ps6, na.rm=T),
            s6 = mean(s6, na.rm=T)) %>%
  mutate(Ratio=ps6/s6)
s6.summary.lane$Norm.Ratio <- s6.summary.lane$Ratio/mean(s6.summary.lane$Ratio[1:4])


#remove outlier lane
#s6.summary.lane <- subset(s6.summary.lane,!( `Lane Name` %in% c("Lane 2","Lane 8")))
kable(s6.summary.lane, caption="Lane Summary for S6", digits=2)

s6.summary <-
  s6.summary.lane %>%
  group_by(Treatment, Diet) %>%
  summarize(Average = mean(Norm.Ratio),
            SE = sd(Norm.Ratio)/sqrt(length(Norm.Ratio)),
            SD = sd(Norm.Ratio),
            n = length(Norm.Ratio))

kable(s6.summary, caption="Summarized Normalized Ratios for S6", digits=2)
s6.plot <-
  s6.summary %>%
  select(Diet,Treatment,Average) %>%
  spread(Treatment, Average)
s6.plot.se <-
  s6.summary %>%
  select(Diet,Treatment,SE) %>%
  spread(Treatment, SE)

ymax <- max(s6.summary$Average + s6.summary$SE)
plot <- barplot(as.matrix(s6.plot[2:3]), beside=T, las=1,
        ylab="pS6/S6", ylim=c(0,ymax))

superpose.eb(plot,
             as.matrix(s6.plot[2:3]),
             as.matrix(s6.plot.se[2:3]))
```

## S6 Statistics

```{r s6-statistics}
s6.aov <- aov(Norm.Ratio~Treatment+Diet, data=s6.summary.lane)
#kable(summary(s6.aov)[[1]], caption="Two Way ANOVA for Effects of Diet and Treatment on Normalized pS6 Ratio", digits=3)
library(car)
```

Ignoring the fed mice, within the fasted mice, the data was normally distributed (Shapiro-Wilk test `r with(s6.summary.lane, shapiro.test(s6.summary.lane[Diet=="CD"&Treatment=="Fasted",]$Norm.Ratio))$p.value` for CD and `r with(s6.summary.lane, shapiro.test(s6.summary.lane[Diet=="HPD"&Treatment=="Fasted",]$Norm.Ratio))$p.value` for HPD), and equal variance could be assumed (Levene's test `r leveneTest(Norm.Ratio~Diet, subset(s6.summary.lane, Treatment=="Fasted"))$"Pr(>F)"[1]`).  Based on this a Student's *t* test had a p-value of `r t.test(Norm.Ratio~Diet, subset(s6.summary.lane, Treatment=="Fasted"), var.equal=T)$p.value`.

# Session Information
```{r sessionInfo, results='asis'}
sessionInfo()
```
