---
title: "Quantification of Blots from McK-TSC1 Knockout Mice Quadriceps"
author: "Dave Bridges"
date: "June 6, 2013"
output:
  html_document:
    keep_md: yes
---

```{r setup, echo=FALSE, cache=FALSE}
## numbers >= 10^5 will be denoted in scientific notation,
## and rounded to 2 digits
options(scipen = 4, digits = 4)
library(knitr)
opts_chunk$set(fig.path='figures/',
               echo=FALSE, warning=FALSE, message=FALSE,dev=c('png','pdf'))
```

# Experimental Details
Sacrificed mice after a 16h fast.  Mice were 69-203 days old. Lysed cells using the Qiagen Tissue Lyser in HNTG buffer (has 1\% Triton X-100).  Blots quantified by ImageJ.

```{r data-entry}
data_file <- "Quadriceps Quantification Data.csv" 
data <- read.csv(data_file)
data$pGS.Ratio <- data$pGS/data$GS
data$GS.PD.Ratio <- data$GS.PD/data$GS
```

# Statistics

```{r data-summary}
require(plyr)
summary <- ddply(
              subset(data, Genotype!="fl/+; Tg/+"),
              ~Genotype, summarise, 
              pGS.mean=mean(pGS.Ratio, na.rm=T),
              GS.PD.mean=mean(GS.PD.Ratio, na.rm=T),
              pGS.se=sd(pGS.Ratio, na.rm=T)/sqrt(length(pGS.Ratio)),
              GS.PD.se=sd(GS.PD.Ratio, na.rm=T)/sqrt(length(GS.PD.Ratio)),
              SREBP.mean = mean(SREBP1c, na.rm=T),
              SREBP.se = sd(SREBP1c, na.rm=T)/sqrt(length(SREBP1c))
              )
summary.normalized <- as.matrix(with(summary, 
                           cbind(pGS.mean/pGS.mean[1],
                                 GS.PD.mean/GS.PD.mean[1],
                                 pGS.se/pGS.mean[1],
                                 GS.PD.se/GS.PD.mean[1],
                                 SREBP.mean/SREBP.mean[1],
                                 SREBP.se/SREBP.mean[1])))
rownames(summary.normalized) <- summary$Genotype
colnames(summary.normalized) <- colnames(summary)[2:7]                           
```

## Effects on GS phosphorylation

A Shapiro-wilk test on the pGS ratio had a p-value of `r min(shapiro.test(subset(data, Genotype=='fl/fl; +/+')$pGS.Ratio)$p.value, shapiro.test(subset(data, Genotype=='fl/fl; Tg/+')$pGS.Ratio)$p.value)`, indicating a non-normal distribution for one of the groups.  Beacause of this we did a Wilcoxon-Rank Sum test, which had a p-value of `r wilcox.test(pGS.Ratio ~ Genotype, data=data)$p.value`.

```{r quadriceps-pGS-ratio}
colors <- c(grey.colors(2)[2], grey.colors(2)[1])
plot <- barplot(summary.normalized[1:2],
        ylab="pGS/GS Ratio (Relative Units)",
        names.arg = summary$Genotype,
        col=colors,
        ylim = c(0,max(summary.normalized[1:2]+summary.normalized[1:2,3])))
superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
  arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
  length = length, ...)
superpose.eb(plot, summary.normalized[1:2], summary.normalized[1:2,3])
legend("topright", c("Wild-Type", "Knockout"), fill=colors, bty="n")
```

```{r quadriceps-GS-PullDown}
plot <- barplot(summary.normalized[1:2,2],
        ylab="GS Ratio (Pull-Down/Total)",
        names.arg = summary$Genotype,
        col=colors,
        ylim = c(0,max(summary.normalized[1:2]+summary.normalized[1:2,4])))
superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
  arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
  length = length, ...)
superpose.eb(plot, summary.normalized[1:2,2], summary.normalized[1:2,4])
legend("topright", c("Wild-Type", "Knockout"), fill=colors, bty="n")

library(car)
```

## Effects on pSREBP1c levels

A Shapiro-wilk test on the processed SREBP1 levels had a p-value of `r min(shapiro.test(subset(data, Genotype=='fl/fl; +/+')$SREBP1c)$p.value, shapiro.test(subset(data, Genotype=='fl/fl; Tg/+')$SREBP1c)$p.value)`, indicating a normal distribution for one of the groups.  We then tested equal variance via Levene's test (p=`r leveneTest(SREBP1c~Genotype, data=data)$"Pr(>F)"[1]`).  Based on this, a Student's *t* test was done (p-value of `r t.test(SREBP1c ~ Genotype, var.equal=T, data=data)$p.value`).

```{r quadriceps-SREBP-levels}
colors <- c(grey.colors(2)[2], grey.colors(2)[1])
plot <- barplot(summary.normalized[9:10],
        ylab="processed SREBP1 (Relative Units)",
        names.arg = summary$Genotype,
        col=colors,
        ylim = c(0,max(summary.normalized[9:10]+summary.normalized[11:12])))

superpose.eb(plot, summary.normalized[9:10], summary.normalized[11:12])
legend("topleft", c("Wild-Type", "Knockout"), fill=colors, bty="n")
```


```{R data-tables}
kable(data, caption = "Raw and Calculated Data", label = "tab:one")
kable(summary.normalized, caption = "Summarized Data", label = "tab:two")
```


# Session Information
```{r session-info}
sessionInfo()
```
