---
title: Longitudinal Body Composition Analysis of Muscle Tsc1 Knockout Mice on a High
  Fat Diet
author: "Erin Stephenson and Dave Bridges"
date: "March 31, 2016"
output:
  html_document:
    keep_md: yes
  pdf_document:
    keep_tex: yes
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.path='figures/',
               echo=FALSE, warning=FALSE, message=FALSE,dev=c('png','pdf'))
options(scipen = 2, digits = 3)


knitr::knit_hooks$set(inline = function(x) {
  knitr:::format_sci(x, 'md')
})
palette( c("blue4","dodgerblue", "deeppink4","pink2"))
```

# Data Entry
This was from combined weights over several measurements of mice all placed on a high fat diet from October 2015-February 2016.  Some animals may appear multiple times in this analysis.  Data is downloaded  from the mousedb website.  This includes only fed weights.  This cohort includes both male and female mice.

```{r data-entry}
raw_data_file <- "Raw Data.csv"
data <- read.csv(raw_data_file, row.names='X')
data$Gender <- relevel(data$animal.Gender, ref="M")
data$Genotype <- relevel(data$animal.Genotype, ref='fl/fl; Tg/+')
data$Knockout <- data$animal.Genotype == 'fl/fl; Tg/+'
data$age.from.start <- data$age - 66
#remove ITT
data <- droplevels(subset(data,experiment.feeding_state=='fed'&assay.assay!='Grip Strength (4 Paw'&assay.assay!='Plasma Glucose'))
data$Weight <- as.numeric(as.character(data$values))/1000

library(dplyr)
library(tidyr)
weight.data <-
  subset(data, age>=67) %>%
  select(assay.assay,animal.id,age,Gender,Genotype,Knockout,Weight) %>%
  distinct(animal.id,assay.assay,age, .keep_all = T) %>%
  spread(assay.assay, Weight) %>%
  mutate(`Percent Fat Mass` = `Total Fat Mass`/`Body Weight`*100)
#convert column names to real names
colnames(weight.data) <- make.names(colnames(weight.data))

#remove outliers
#NEED TO REMOVE DATA FROM WEEK 2 OF HFD- MICE WERE GIVEN A POST-CLAMS RECOVERY WEEK AND THEREFORE NOT MANY WERE MEASURED DURING THIS TIMEPOINT


#mouse 4197 died in cage of unknown causes and was partially eaten by cagemates very early into HFD enrollment
weight.data <-subset(weight.data, !(animal.id==9219))
#mouse 4508 was likely sick (lost weight rapidly and died very early into HFD enrollment)
weight.data <- subset(weight.data, !(animal.id==9503))
#removed 4470 as it had mangled ears and was euthanized early on in the study
weight.data <- subset(weight.data, !(animal.id==9494))


se <- function(x) sd(x, na.rm=T)/sqrt(length(x))
combined.summary <-
  subset(weight.data, age>=67) %>%
  mutate(Weeks = round((age-67)/7)) %>%
  group_by(Weeks,Gender,Knockout) %>%
  distinct(Weeks,animal.id, .keep_all = T) %>%
  summarise(Body.Weight.mean = mean(Body.Weight, na.rm=T),
            Body.Weight.se = se(Body.Weight),
            Total.Fat.Mass.mean = mean(Total.Fat.Mass, na.rm=T),
            Total.Fat.Mass.se = se(Total.Fat.Mass),
            Lean.Mass.mean = mean(Lean.Mass, na.rm=T),
            Lean.Mass.se = se(Lean.Mass),
            Percent.Fat.Mass.mean = mean(Percent.Fat.Mass, na.rm=T),
            Percent.Fat.Mass.se = se(Percent.Fat.Mass))

combined.summary <- subset(combined.summary, Weeks!=1)
```

Data was downloaded from MouseDB then aand the data is saved as `r raw_data_file`.  These data are located in `r getwd()` and was most recently updated on `r date()`.

# Enrollment

```{r enrollment}
counts.genotype <- 
  weight.data %>%
  group_by(Genotype,Gender) %>%
  distinct(animal.id, .keep_all = T) %>%
  summarize(Number = length(age))
kable(counts.genotype, caption="Total Animals Enrolled in Study by Genotype")

counts.knockout <- 
  weight.data %>%
  group_by(Knockout,Gender) %>%
  distinct(animal.id, .keep_all = T) %>%
  summarize(Number = length(age))
kable(counts.knockout, caption="Total Animals Enrolled in Study by Knockout Status")
```

# Body Weights

```{r weights-scatterplot}
ymin = min(weight.data$Body.Weight, na.rm=T)
ymax = max(weight.data$Body.Weight, na.rm=T)

with(subset(weight.data, Knockout == TRUE&Gender==levels(Gender)[1]), 
     plot(age, Body.Weight,
                       col=palette()[1],
                       ylab="Body Weight (g)",
                       xlab='Age (days)',
                       ylim=c(ymin,ymax),
                        las=1, pch=19))
with(subset(weight.data, Knockout == FALSE&Gender==levels(Gender)[1]), 
     points(age, Body.Weight, col=palette()[2], pch=19))
with(subset(weight.data, Knockout == TRUE&Gender==levels(Gender)[2]), 
     points(age, Body.Weight, col=palette()[3], pch=19))
with(subset(weight.data, Knockout == FALSE&Gender==levels(Gender)[2]), 
     points(age, Body.Weight, col=palette()[4], pch=19))

lines(with(subset(weight.data, Gender==levels(Gender)[1]&Knockout == TRUE), 
           loess.smooth(age,Body.Weight)), col=palette()[1])
lines(with(subset(weight.data, Gender==levels(Gender)[1]&Knockout == FALSE), 
           loess.smooth(age,Body.Weight)), col=palette()[2])
lines(with(subset(weight.data, Gender==levels(Gender)[2]&Knockout == TRUE), 
           loess.smooth(age,Body.Weight)), col=palette()[3])
lines(with(subset(weight.data, Gender==levels(Gender)[2]&Knockout == FALSE), 
           loess.smooth(age,Body.Weight)), col=palette()[4])


legend("topleft", do.call(paste, expand.grid(c("Knockout","Wild-Type"),'-',levels(weight.data$Gender))),
       lty=1, col=palette()[1:4], bty='n', cex=1)

yaxes <- with(combined.summary, 
             c(min(Body.Weight.mean, na.rm=T)-max(Body.Weight.se, na.rm=T), 
             c(max(Body.Weight.mean, na.rm=T)+max(Body.Weight.se, na.rm=T))))
plot <- with(subset(combined.summary, 
                   Gender==levels(Gender)[1]&Knockout == TRUE), 
             plot(Weeks, Body.Weight.mean, type="l", ylim=yaxes,
                  las=1, ylab="Body Weight (g)", col=palette()[1]))
with(subset(combined.summary, 
                    Gender==levels(Gender)[1]&Knockout == FALSE), 
             lines(Weeks, Body.Weight.mean, col=palette()[2]))
with(subset(combined.summary, 
                    Gender==levels(Gender)[2]&Knockout == TRUE), 
             lines(Weeks, Body.Weight.mean, col=palette()[3]))
with(subset(combined.summary, 
                    Gender==levels(Gender)[2]&Knockout == FALSE), 
             lines(Weeks, Body.Weight.mean, col=palette()[4]))

superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
  arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
  length = length, ...)

with(subset(combined.summary, 
                  Gender==levels(Gender)[1]&Knockout == TRUE), 
            superpose.eb(Weeks, Body.Weight.mean, Body.Weight.se, col=palette()[1]))
with(subset(combined.summary, 
                   Gender==levels(Gender)[1]&Knockout == FALSE), 
             superpose.eb(Weeks, Body.Weight.mean, Body.Weight.se, col=palette()[2]))
with(subset(combined.summary, 
                   Gender==levels(Gender)[2]&Knockout == TRUE), 
             superpose.eb(Weeks, Body.Weight.mean, Body.Weight.se, col=palette()[3]))
with(subset(combined.summary, 
                   Gender==levels(Gender)[2]&Knockout == FALSE), 
             superpose.eb(Weeks, Body.Weight.mean, Body.Weight.se, col=palette()[4]))

legend("topleft", do.call(paste, expand.grid(c("Knockout","Wild-Type"),'-',levels(weight.data$Gender))),
       lty=1, col=palette()[1:4], bty='n', cex=1)

#only up to 12w
yaxes <- with(subset(combined.summary,Weeks<12), 
             c(min(Body.Weight.mean, na.rm=T)-max(Body.Weight.se, na.rm=T), 
             c(max(Body.Weight.mean, na.rm=T)+max(Body.Weight.se, na.rm=T))))
plot <- with(subset(combined.summary, 
                   Gender==levels(Gender)[1]&Knockout == TRUE&Weeks<12), 
             plot(Weeks, Body.Weight.mean, type="l", ylim=yaxes,
                  las=1, ylab="Body Weight (g)", col=palette()[1]))
with(subset(combined.summary, 
                    Gender==levels(Gender)[1]&Knockout == FALSE&Weeks<12), 
             lines(Weeks, Body.Weight.mean, col=palette()[2]))
with(subset(combined.summary, 
                    Gender==levels(Gender)[2]&Knockout == TRUE&Weeks<12), 
             lines(Weeks, Body.Weight.mean, col=palette()[3]))
with(subset(combined.summary, 
                    Gender==levels(Gender)[2]&Knockout == FALSE&Weeks<12), 
             lines(Weeks, Body.Weight.mean, col=palette()[4]))

with(subset(combined.summary, 
                  Gender==levels(Gender)[1]&Knockout == TRUE&Weeks<12), 
            superpose.eb(Weeks, Body.Weight.mean, Body.Weight.se, col=palette()[1]))
with(subset(combined.summary, 
                   Gender==levels(Gender)[1]&Knockout == FALSE&Weeks<12), 
             superpose.eb(Weeks, Body.Weight.mean, Body.Weight.se, col=palette()[2]))
with(subset(combined.summary, 
                   Gender==levels(Gender)[2]&Knockout == TRUE&Weeks<12), 
             superpose.eb(Weeks, Body.Weight.mean, Body.Weight.se, col=palette()[3]))
with(subset(combined.summary, 
                   Gender==levels(Gender)[2]&Knockout == FALSE&Weeks<12), 
             superpose.eb(Weeks, Body.Weight.mean, Body.Weight.se, col=palette()[4]))

legend("topleft", do.call(paste, expand.grid(c("Knockout","Wild-Type"),'-',levels(weight.data$Gender))),
       lty=1, col=palette()[1:4], bty='n', cex=1)

#statistics
library(lme4)
weight.lme <- lmer(Body.Weight~age*Gender*Knockout + (1|animal.id) + (age-1|animal.id), data=subset(weight.data,age<12*7+70), REML=F)
weight.lme.null <- lmer(Body.Weight~age*Gender + (1|animal.id) + (age-1|animal.id), data=subset(weight.data,age<12*7+70), REML=F)

#wt.predict.m <- predict(weight.lme, newdata=list(age=7*12+70, Gender="M", Knockout=FALSE, animal.id=weight.data$animal.id[1]))
#ko.predict.m <- predict(weight.lme, newdata=list(age=7*12+70, Gender="M", Knockout=TRUE, animal.id=weight.data$animal.id[1]))

#wt.predict.f <- predict(weight.lme, newdata=list(age=7*12+70, Gender="F", Knockout=FALSE, animal.id=weight.data$animal.id[1]))
#ko.predict.f <- predict(weight.lme, newdata=list(age=7*12+70, Gender="F", Knockout=TRUE, animal.id=weight.data$animal.id[1]))

#fixef(weight.lme) %>% kable(caption="Coefficients for body weight model")
```

To test whether these groups are different we constructed a linear model with the following formula:

Weight ~ `r as.character(formula(weight.lme))[3]`

The knockout caused a s (p=`r anova(weight.lme, weight.lme.null)$"Pr(>Chisq)"[2]` via a Chi-squared test).

The full results are shown below:

```{r weight-lmer-results}
coefs <- data.frame(coef(summary(weight.lme)))
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
kable(coefs, caption="Estimates and p-values from mixed linear models.")
```



```{r diagnostics-body-weight}
#statistics
library(lme4)
weight.cd.hpd.lme <- lmer(Body.Weight ~ age * Gender * Knockout + (1|animal.id) + (age-1|animal.id), data=weight.data, REML=F)
weight.cd.hpd.lme.null <- lmer(Body.Weight ~ age * Gender + (1|animal.id) + (age-1|animal.id), data=weight.data, REML=F)
#using 
library(influence.ME)
infl <- influence(weight.cd.hpd.lme, group='animal.id')
par(mfrow=c(1,3))
#plot(density(residuals(weight.cd.hpd.lme)), main="Effects of High Protein Knockout", las=1)
#plot(fitted(weight.cd.hpd.lme),residuals(weight.cd.hpd.lme),  ylab="Residuals", xlab="Fitted Values (g)", pch=19, las=1)
#abline(h=0, lty=2, col='red')
#barplot(cooks.distance(infl), las=2, names.arg=rownames(cooks.distance(infl)),  beside=T, ylab="Cook's Distance", col='grey')
```

# Lean Mass

```{r lean-mass-scatterplot}
ymin = min(weight.data$Lean.Mass, na.rm=T)
ymax = max(weight.data$Lean.Mass, na.rm=T)

with(subset(weight.data, Gender==levels(Gender)[1]&Knockout == TRUE)
     , plot(age, Lean.Mass,
                       col=palette()[1],
                      ylim=c(ymin,ymax),
                       ylab="Total Lean Mass (g)",
                       xlab='Age (days)',
                        las=1, pch=19))
with(subset(weight.data, Gender==levels(Gender)[1]&Knockout == FALSE), 
     points(age, Lean.Mass, col=palette()[2], pch=19))
with(subset(weight.data, Gender==levels(Gender)[2]&Knockout == TRUE), 
     points(age, Lean.Mass, col=palette()[3], pch=19))
with(subset(weight.data, Gender==levels(Gender)[2]&Knockout == FALSE), 
     points(age, Lean.Mass, col=palette()[4], pch=19))

lines(with(subset(weight.data, Gender==levels(Gender)[1]&Knockout == TRUE), 
           loess.smooth(age,Lean.Mass)), col=palette()[1])
lines(with(subset(weight.data, Gender==levels(Gender)[1]&Knockout == FALSE), 
           loess.smooth(age,Lean.Mass)), col=palette()[2])
lines(with(subset(weight.data, Gender==levels(Gender)[2]&Knockout == TRUE), 
           loess.smooth(age,Lean.Mass)), col=palette()[3])
lines(with(subset(weight.data, Gender==levels(Gender)[2]&Knockout == FALSE), 
           loess.smooth(age,Lean.Mass)), col=palette()[4])


legend("topleft", do.call(paste, expand.grid(c("Knockout","Wild-Type"),'-',levels(weight.data$Gender))),
       lty=1, col=palette()[1:4], bty='n', cex=1)

yaxes <- with(combined.summary, 
              c(min(Lean.Mass.mean, na.rm=T)-max(Lean.Mass.se, na.rm=T), 
               c(max(Lean.Mass.mean, na.rm=T)+max(Lean.Mass.se, na.rm=T))))
plot <- with(subset(combined.summary, 
                    Gender==levels(Gender)[1]&Knockout == TRUE), 
             plot(Weeks, Lean.Mass.mean, type="l", ylim=yaxes,
                  las=1, ylab="Total Lean Mass (g)", col=palette()[1]))
with(subset(combined.summary, 
                    Gender==levels(Gender)[1]&Knockout == FALSE), 
             lines(Weeks, Lean.Mass.mean, col=palette()[2]))
with(subset(combined.summary, 
                    Gender==levels(Gender)[2]&Knockout == TRUE), 
             lines(Weeks, Lean.Mass.mean, col=palette()[3]))
with(subset(combined.summary, 
                    Gender==levels(Gender)[2]&Knockout == FALSE), 
             lines(Weeks, Lean.Mass.mean, col=palette()[4]))

with(subset(combined.summary, 
                  Gender==levels(Gender)[1]&Knockout == TRUE), 
            superpose.eb(Weeks, Lean.Mass.mean, Lean.Mass.se, col=palette()[1]))
with(subset(combined.summary, 
                   Gender==levels(Gender)[1]&Knockout == FALSE), 
             superpose.eb(Weeks, Lean.Mass.mean, Lean.Mass.se, col=palette()[2]))
with(subset(combined.summary, 
                   Gender==levels(Gender)[2]&Knockout == TRUE), 
             superpose.eb(Weeks, Lean.Mass.mean, Lean.Mass.se, col=palette()[3]))
with(subset(combined.summary, 
                   Gender==levels(Gender)[2]&Knockout == FALSE), 
             superpose.eb(Weeks, Lean.Mass.mean, Lean.Mass.se, col=palette()[4]))

legend("topleft", do.call(paste, expand.grid(c("Knockout","Wild-Type"),'-',levels(weight.data$Gender))),
       lty=1, col=palette()[1:4], bty='n', cex=1)

#up to 12 w only

yaxes <- with(subset(combined.summary, Weeks<12),
              c(min(Lean.Mass.mean, na.rm=T)-max(Lean.Mass.se, na.rm=T), 
               c(max(Lean.Mass.mean, na.rm=T)+max(Lean.Mass.se, na.rm=T))))
plot <- with(subset(combined.summary, 
                    Gender==levels(Gender)[1]&Knockout == TRUE&Weeks<12), 
             plot(Weeks, Lean.Mass.mean, type="l", ylim=yaxes,
                  las=1, ylab="Total Lean Mass (g)", col=palette()[1]))
with(subset(combined.summary, 
                    Gender==levels(Gender)[1]&Knockout == FALSE&Weeks<12), 
             lines(Weeks, Lean.Mass.mean, col=palette()[2]))
with(subset(combined.summary, 
                    Gender==levels(Gender)[2]&Knockout == TRUE&Weeks<12), 
             lines(Weeks, Lean.Mass.mean, col=palette()[3]))
with(subset(combined.summary, 
                    Gender==levels(Gender)[2]&Knockout == FALSE&Weeks<12), 
             lines(Weeks, Lean.Mass.mean, col=palette()[4]))

with(subset(combined.summary, 
                  Gender==levels(Gender)[1]&Knockout == TRUE&Weeks<12), 
            superpose.eb(Weeks, Lean.Mass.mean, Lean.Mass.se, col=palette()[1]))
with(subset(combined.summary, 
                   Gender==levels(Gender)[1]&Knockout == FALSE&Weeks<12), 
             superpose.eb(Weeks, Lean.Mass.mean, Lean.Mass.se, col=palette()[2]))
with(subset(combined.summary, 
                   Gender==levels(Gender)[2]&Knockout == TRUE&Weeks<12), 
             superpose.eb(Weeks, Lean.Mass.mean, Lean.Mass.se, col=palette()[3]))
with(subset(combined.summary, 
                   Gender==levels(Gender)[2]&Knockout == FALSE&Weeks<12), 
             superpose.eb(Weeks, Lean.Mass.mean, Lean.Mass.se, col=palette()[4]))

legend("topleft", do.call(paste, expand.grid(c("Knockout","Wild-Type"),'-',levels(weight.data$Gender))),
       lty=1, col=palette()[1:4], bty='n', cex=1)
```

## Lean Mass Statistics

```{r lean-mass-stats}
#base model
lean.lme.base <- lmer(Lean.Mass~age + (1|animal.id) + (age-1|animal.id), data=subset(weight.data,age<12*7+70), REML=F)

#is there an effect of sex
lean.lme.sex <- lmer(Lean.Mass~age + Gender + age:Gender+ (1|animal.id) + (age-1|animal.id), data=subset(weight.data,age<12*7+70), REML=F)

#is there an effect of knockout (adding interaction with age to account for slope effects)
lean.lme.sex.ko <- lmer(Lean.Mass~age + 
                         Gender + 
                         age:Gender + 
                         Knockout + 
                         Knockout:age + 
                         (1|animal.id) + (age-1|animal.id), 
                       data=subset(weight.data,age<12*7+70), REML=F)

#is there an interaction of knockout with sex
lean.lme.sex.ko.int <- lmer(Lean.Mass~age + 
                             Gender + 
                             age:Gender + 
                             Knockout + 
                             Knockout:age + 
                             Gender:Knockout + 
                             Gender:age:Knockout + 
                             (1|animal.id) + (age-1|animal.id), data=subset(weight.data,age<12*7+70), REML=F)


lean.lme.null <- lmer(Lean.Mass~age + Gender + age:Gender + (1|animal.id) + (age-1|animal.id), data=subset(weight.data,age<12*7+70), REML=F)

```

To test whether these groups are different we constructed a linear model with the following formula:

Fat-Free Mass ~ `r as.character(formula(lean.lme.sex.ko.int))[3]`.  

We used this model because the base model was that Lean Mass changes with age.  We asked if sex modified the age dependent effect, and it did (p=`r anova(lean.lme.base, lean.lme.sex)$"Pr(>Chisq)"[2]`).  We next added knockout to the modified sex-modified model, by adding both a knockout and knockout:age term and it was not significant (p=`r anova(lean.lme.sex, lean.lme.sex.ko)$"Pr(>Chisq)"[2]`).  


Based on this formula the knockout caused significant reductions in lean mass, via a Chi-squared test with a model not including genotype (p=`r anova(lean.lme.sex.ko, lean.lme.null)$"Pr(>Chisq)"[2]`).

The full results are shown below:

```{r lean-lmer-results}
coefs <- data.frame(coef(summary(lean.lme.sex.ko)))
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
kable(coefs, caption="Estimates and p-values from mixed linear models.")
```

# Fat Mass

```{r fat-mass-scatterplot}
with(subset(weight.data, Knockout == TRUE&Gender==levels(Gender)[1]), 
     plot(age, Total.Fat.Mass,ylim=c(0,15),
                       col=palette()[1],
                       ylab="Total Fat Mass (g)",
                       xlab='Age (days)',
                        las=1, pch=19))
with(subset(weight.data, Knockout == FALSE&Gender==levels(Gender)[1]), 
     points(age, Total.Fat.Mass, col=palette()[2], pch=19))
with(subset(weight.data, Knockout == TRUE&Gender==levels(Gender)[2]), 
     points(age, Total.Fat.Mass, col=palette()[3], pch=19))
with(subset(weight.data, Knockout == FALSE&Gender==levels(Gender)[2]), 
     points(age, Total.Fat.Mass, col=palette()[4], pch=19))

lines(with(subset(weight.data, Gender==levels(Gender)[1]&Knockout == TRUE), 
           loess.smooth(age,Total.Fat.Mass)), col=palette()[1])
lines(with(subset(weight.data, Gender==levels(Gender)[1]&Knockout == FALSE), 
           loess.smooth(age,Total.Fat.Mass)), col=palette()[2])
lines(with(subset(weight.data, Gender==levels(Gender)[2]&Knockout == TRUE), 
           loess.smooth(age,Total.Fat.Mass)), col=palette()[3])
lines(with(subset(weight.data, Gender==levels(Gender)[2]&Knockout == FALSE), 
           loess.smooth(age,Total.Fat.Mass)), col=palette()[4])


legend("topleft", do.call(paste, expand.grid(c("Knockout","Wild-Type"),'-',levels(weight.data$Gender))),
       lty=1, col=palette()[1:4], bty='n', cex=1)

yaxes <- with(combined.summary, 
              c(min(Total.Fat.Mass.mean, na.rm=T)-max(Total.Fat.Mass.se, na.rm=T), 
               c(max(Total.Fat.Mass.mean, na.rm=T)+max(Total.Fat.Mass.se, na.rm=T))))
plot <- with(subset(combined.summary, 
             Gender==levels(Gender)[1]&Knockout == TRUE),
             plot(Weeks, Total.Fat.Mass.mean, type="l", ylim=yaxes,
                  las=1, ylab="Total Fat Mass (g)", col=palette()[1]))
with(subset(combined.summary, 
             Gender==levels(Gender)[1]&Knockout == FALSE),
             lines(Weeks, Total.Fat.Mass.mean, col=palette()[2]))
with(subset(combined.summary, 
             Gender==levels(Gender)[2]&Knockout == TRUE),
             lines(Weeks, Total.Fat.Mass.mean, col=palette()[3]))
with(subset(combined.summary, 
             Gender==levels(Gender)[2]&Knockout == FALSE),
           lines(Weeks, Total.Fat.Mass.mean, col=palette()[4]))

with(subset(combined.summary, 
             Gender==levels(Gender)[1]&Knockout == TRUE),
             superpose.eb(Weeks, Total.Fat.Mass.mean, Total.Fat.Mass.se, col=palette()[1]))
with(subset(combined.summary, 
             Gender==levels(Gender)[1]&Knockout == FALSE),
            superpose.eb(Weeks, Total.Fat.Mass.mean, Total.Fat.Mass.se, col=palette()[2]))
with(subset(combined.summary, 
             Gender==levels(Gender)[2]&Knockout == TRUE),
             superpose.eb(Weeks, Total.Fat.Mass.mean, Total.Fat.Mass.se, col=palette()[3]))
with(subset(combined.summary, 
             Gender==levels(Gender)[2]&Knockout == FALSE),
             superpose.eb(Weeks, Total.Fat.Mass.mean, Total.Fat.Mass.se, col=palette()[4]))

legend("topleft", do.call(paste, expand.grid(c("Knockout","Wild-Type"),'-',levels(weight.data$Gender))),
       lty=1, col=palette()[1:4], bty='n', cex=1)

#after 12w only

yaxes <- with(subset(combined.summary, Weeks<12),
              c(min(Total.Fat.Mass.mean, na.rm=T)-max(Total.Fat.Mass.se, na.rm=T), 
               c(max(Total.Fat.Mass.mean, na.rm=T)+max(Total.Fat.Mass.se, na.rm=T))))
plot <- with(subset(combined.summary, 
             Gender==levels(Gender)[1]&Knockout == TRUE& Weeks<12),
             plot(Weeks, Total.Fat.Mass.mean, type="l", ylim=yaxes,
                  las=1, ylab="Total Fat Mass (g)", col=palette()[1]))
with(subset(combined.summary, 
             Gender==levels(Gender)[1]&Knockout == FALSE& Weeks<12),
             lines(Weeks, Total.Fat.Mass.mean, col=palette()[2]))
with(subset(combined.summary, 
             Gender==levels(Gender)[2]&Knockout == TRUE& Weeks<12),
             lines(Weeks, Total.Fat.Mass.mean, col=palette()[3]))
with(subset(combined.summary, 
             Gender==levels(Gender)[2]&Knockout == FALSE& Weeks<12),
           lines(Weeks, Total.Fat.Mass.mean, col=palette()[4]))

with(subset(combined.summary, 
             Gender==levels(Gender)[1]&Knockout == TRUE& Weeks<12),
             superpose.eb(Weeks, Total.Fat.Mass.mean, Total.Fat.Mass.se, col=palette()[1]))
with(subset(combined.summary, 
             Gender==levels(Gender)[1]&Knockout == FALSE& Weeks<12),
            superpose.eb(Weeks, Total.Fat.Mass.mean, Total.Fat.Mass.se, col=palette()[2]))
with(subset(combined.summary, 
             Gender==levels(Gender)[2]&Knockout == TRUE& Weeks<12),
             superpose.eb(Weeks, Total.Fat.Mass.mean, Total.Fat.Mass.se, col=palette()[3]))
with(subset(combined.summary, 
             Gender==levels(Gender)[2]&Knockout == FALSE& Weeks<12),
             superpose.eb(Weeks, Total.Fat.Mass.mean, Total.Fat.Mass.se, col=palette()[4]))

legend("topleft", do.call(paste, expand.grid(c("Knockout","Wild-Type"),'-',levels(weight.data$Gender))),
       lty=1, col=palette()[1:4], bty='n', cex=1)
```

## Statistics for Fat Mass 

```{r statistics-fat-mass}
#base model
fat.lme.base <- lmer(Total.Fat.Mass~age + (1|animal.id) + (age-1|animal.id), data=subset(weight.data,age<12*7+70), REML=F)

#is there an effect of sex
fat.lme.sex <- lmer(Total.Fat.Mass~age + Gender + age:Gender+ (1|animal.id) + (age-1|animal.id), data=subset(weight.data,age<12*7+70), REML=F)

#is there an effect of knockout (adding interaction with age to account for slope effects)
fat.lme.sex.ko <- lmer(Total.Fat.Mass~age + 
                         Gender + 
                         age:Gender + 
                         Knockout + 
                         Knockout:age + 
                         (1|animal.id) + (age-1|animal.id), 
                       data=subset(weight.data,age<12*7+70), REML=F)

#is there an interaction of knockout with sex
fat.lme.sex.ko.int <- lmer(Total.Fat.Mass~age + 
                             Gender + 
                             age:Gender + 
                             Knockout + 
                             Knockout:age + 
                             Gender:Knockout + 
                             Gender:age:Knockout + 
                             (1|animal.id) + (age-1|animal.id), data=subset(weight.data,age<12*7+70), REML=F)


fat.lme.null <- lmer(Total.Fat.Mass~age + Gender + age:Gender + (1|animal.id) + (age-1|animal.id), data=subset(weight.data,age<12*7+70), REML=F)
```

To test whether these groups are different we constructed a linear model with the following formula:

Fat Mass ~ `r as.character(formula(fat.lme.sex.ko.int))[3]`.  

We used this model because the base model was that Fat Mass changes with age.  We asked if sex modified the age dependent effect, and it did (p=`r anova(fat.lme.base, fat.lme.sex)$"Pr(>Chisq)"[2]`).  We next added knockout to the modified sex-modified model, by adding both a knockout and knockout:age term and it too was significant (p=`r anova(fat.lme.sex, fat.lme.sex.ko)$"Pr(>Chisq)"[2]`).  

Finally we asked if sex modified that last model by adding Gender:Knockout and Gender:age:Knockout terms.  This ANOVA yielded a p-value of `r anova(fat.lme.sex.ko, fat.lme.sex.ko.int)$"Pr(>Chisq)"[2]`, which was also significant.

Based on this formula the knockout caused significant reductions in mass, via a Chi-squared test with a model not including genotype (p=`r anova(fat.lme.sex.ko.int, fat.lme.null)$"Pr(>Chisq)"[2]`).

The full results are shown below:

```{r fat-lmer-results}
coefs <- data.frame(coef(summary(fat.lme.sex.ko.int)))
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
kable(coefs, caption="Estimates and p-values from mixed linear models.")
```


# Percent Fat Mass

```{r percent-fat-mass-scatterplot}
with(subset(weight.data, Knockout == TRUE&Gender==levels(Gender)[1]), 
plot(age, Percent.Fat.Mass,ylim=c(0,35),
                       col=Gender,
                       ylab="Percent Fat Mass",
                       xlab='Age (days)',
                        las=1, pch=19))

with(subset(weight.data, Knockout == FALSE&Gender==levels(Gender)[1]), 
     points(age, Percent.Fat.Mass, col=palette()[2], pch=19))
with(subset(weight.data, Knockout == TRUE&Gender==levels(Gender)[2]), 
     points(age, Percent.Fat.Mass, col=palette()[3], pch=19))
with(subset(weight.data, Knockout == FALSE&Gender==levels(Gender)[2]), 
     points(age, Percent.Fat.Mass, col=palette()[4], pch=19))

lines(with(subset(weight.data, Gender==levels(Gender)[1]&Knockout == TRUE), 
           loess.smooth(age,Percent.Fat.Mass)), col=palette()[1])
lines(with(subset(weight.data, Gender==levels(Gender)[1]&Knockout == FALSE), 
           loess.smooth(age,Percent.Fat.Mass)), col=palette()[2])
lines(with(subset(weight.data, Gender==levels(Gender)[2]&Knockout == TRUE), 
           loess.smooth(age,Percent.Fat.Mass)), col=palette()[3])
lines(with(subset(weight.data, Gender==levels(Gender)[2]&Knockout == FALSE), 
           loess.smooth(age,Percent.Fat.Mass)), col=palette()[4])

legend("topleft", do.call(paste, expand.grid(c("Knockout","Wild-Type"),'-',levels(weight.data$Gender))),
       lty=1, col=palette()[1:4], bty='n', cex=1)

#percent fat mass with error bars
yaxes <- with(combined.summary, 
              c(min(Percent.Fat.Mass.mean, na.rm=T)-max(Percent.Fat.Mass.se, na.rm=T), 
               c(max(Percent.Fat.Mass.mean, na.rm=T)+max(Percent.Fat.Mass.se, na.rm=T))))
plot <- with(subset(combined.summary, 
                    Gender==levels(Gender)[1]&Knockout == TRUE), 
             plot(Weeks, Percent.Fat.Mass.mean, type="l", ylim=yaxes,
                  las=1, ylab="Percent Fat Mass", col=palette()[1]))
with(subset(combined.summary, 
                    Gender==levels(Gender)[1]&Knockout == FALSE), 
             lines(Weeks, Percent.Fat.Mass.mean, col=palette()[2]))
with(subset(combined.summary, 
                    Gender==levels(Gender)[2]&Knockout == TRUE), 
             lines(Weeks, Percent.Fat.Mass.mean, col=palette()[3]))
with(subset(combined.summary, 
                    Gender==levels(Gender)[2]&Knockout == FALSE), 
             lines(Weeks, Percent.Fat.Mass.mean, col=palette()[4]))

with(subset(combined.summary, 
                    Gender=="M"&Knockout=="Control Knockout"), 
             superpose.eb(Weeks, Percent.Fat.Mass.mean, Percent.Fat.Mass.se, col=palette()[1]))
with(subset(combined.summary, 
                    Gender=="M"&Knockout=="High Protein Knockout"), 
             superpose.eb(Weeks, Percent.Fat.Mass.mean, Percent.Fat.Mass.se, col=palette()[2]))
with(subset(combined.summary, 
                    Gender=="F"&Knockout=="Control Knockout"), 
             superpose.eb(Weeks, Percent.Fat.Mass.mean, Percent.Fat.Mass.se, col=palette()[3]))
with(subset(combined.summary, 
                    Gender=="F"&Knockout=="High Protein Knockout"), 
             superpose.eb(Weeks, Percent.Fat.Mass.mean, Percent.Fat.Mass.se, col=palette()[4]))

with(subset(combined.summary, 
                  Gender==levels(Gender)[1]&Knockout == TRUE), 
            superpose.eb(Weeks, Percent.Fat.Mass.mean, Percent.Fat.Mass.se, col=palette()[1]))
with(subset(combined.summary, 
                   Gender==levels(Gender)[1]&Knockout == FALSE), 
             superpose.eb(Weeks, Percent.Fat.Mass.mean, Percent.Fat.Mass.se, col=palette()[2]))
with(subset(combined.summary, 
                   Gender==levels(Gender)[2]&Knockout == TRUE), 
             superpose.eb(Weeks, Percent.Fat.Mass.mean, Percent.Fat.Mass.se, col=palette()[3]))
with(subset(combined.summary, 
                   Gender==levels(Gender)[2]&Knockout == FALSE), 
             superpose.eb(Weeks, Percent.Fat.Mass.mean, Percent.Fat.Mass.se, col=palette()[4]))

legend("topleft", do.call(paste, expand.grid(c("Knockout","Wild-Type"),'-',levels(weight.data$Gender))),
       lty=1, col=palette()[1:4], bty='n', cex=1)

#up to 12w only

yaxes <- with(subset(combined.summary, Weeks<12),
              c(min(Percent.Fat.Mass.mean, na.rm=T)-max(Percent.Fat.Mass.se, na.rm=T), 
               c(max(Percent.Fat.Mass.mean, na.rm=T)+max(Percent.Fat.Mass.se, na.rm=T))))
plot <- with(subset(combined.summary, 
                    Gender==levels(Gender)[1]&Knockout == TRUE & Weeks<12), 
             plot(Weeks, Percent.Fat.Mass.mean, type="l", ylim=yaxes,
                  las=1, ylab="Percent Fat Mass", col=palette()[1]))
with(subset(combined.summary, 
                    Gender==levels(Gender)[1]&Knockout == FALSE & Weeks<12), 
             lines(Weeks, Percent.Fat.Mass.mean, col=palette()[2]))
with(subset(combined.summary, 
                    Gender==levels(Gender)[2]&Knockout == TRUE & Weeks<12), 
             lines(Weeks, Percent.Fat.Mass.mean, col=palette()[3]))
with(subset(combined.summary, 
                    Gender==levels(Gender)[2]&Knockout == FALSE& Weeks<12), 
             lines(Weeks, Percent.Fat.Mass.mean, col=palette()[4]))

with(subset(combined.summary, 
                  Gender==levels(Gender)[1]&Knockout == TRUE& Weeks<12), 
            superpose.eb(Weeks, Percent.Fat.Mass.mean, Percent.Fat.Mass.se, col=palette()[1]))
with(subset(combined.summary, 
                   Gender==levels(Gender)[1]&Knockout == FALSE& Weeks<12), 
             superpose.eb(Weeks, Percent.Fat.Mass.mean, Percent.Fat.Mass.se, col=palette()[2]))
with(subset(combined.summary, 
                   Gender==levels(Gender)[2]&Knockout == TRUE& Weeks<12), 
             superpose.eb(Weeks, Percent.Fat.Mass.mean, Percent.Fat.Mass.se, col=palette()[3]))
with(subset(combined.summary, 
                   Gender==levels(Gender)[2]&Knockout == FALSE& Weeks<12), 
             superpose.eb(Weeks, Percent.Fat.Mass.mean, Percent.Fat.Mass.se, col=palette()[4]))

legend("topleft", do.call(paste, expand.grid(c("Knockout","Wild-Type"),'-',levels(weight.data$Gender))),
       lty=1, col=palette()[1:4], bty='n', cex=1)


pct.fat.lme <- lmer(Percent.Fat.Mass~age*Gender*Knockout + (1|animal.id) + (age-1|animal.id), data=subset(weight.data,age<12*7+70), REML=F)
pct.fat.lme.null <- lmer(Percent.Fat.Mass~age*Gender + (1|animal.id) + (age-1|animal.id), data=subset(weight.data,age<12*7+70), REML=F)

#wt.predict.m <- predict(pct.fat.lme, newdata=list(age=7*12+70, Gender="M", Knockout=FALSE, animal.id=weight.data$animal.id[1]))
#ko.predict.m <- predict(pct.fat.lme, newdata=list(age=7*12+70, Gender="M", Knockout=TRUE, animal.id=weight.data$animal.id[1]))

#wt.predict.f <- predict(pct.fat.lme, newdata=list(age=7*12+70, Gender="F", Knockout=FALSE, animal.id=weight.data$animal.id[1]))
#ko.predict.f <- predict(pct.fat.lme, newdata=list(age=7*12+70, Gender="F", Knockout=TRUE, animal.id=weight.data$animal.id[1]))

```

To test whether these groups are different we constructed a linear model with the following formula:

Percent Fat Mass~ `r as.character(formula(pct.fat.lme))[3]`

Based on this formula the knockout caused a increaes for females (p=`r anova(pct.fat.lme, pct.fat.lme.null)$"Pr(>Chisq)"[2]` via a Chi-squared test).

The full results are shown below:

```{r pct-fat-results}
coefs <- data.frame(coef(summary(pct.fat.lme)))
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
kable(coefs, caption="Estimates and p-values from mixed linear models.")
```

# End of Study Summary

The following animals have completed the study (age is >`r  67+7*12-7` days):

```{r end-of-study}
weight.data %>%
   subset(age >= 67+7*12-7) %>%
   group_by(Genotype,Gender) %>%
   distinct(animal.id, .keep_all = T) %>%
   summarize(Number = length(age)) %>%
  kable(caption="Animals which have completed the study by genotype")

weight.data %>%
  subset(age >= 67+7*12-7) %>%
  group_by(Knockout,Gender) %>%
  distinct(animal.id, .keep_all = T) %>%
  summarize(Number = length(age)) %>%
kable(caption="Animals which have completed the study, by knockout status")

summary.end <- 
  subset(weight.data, age >= 67+7*12-7) %>%
  group_by(Gender, Knockout) %>%
  distinct(animal.id, .keep_all = T) %>%
  select(-Genotype, -animal.id, -age) %>%
  summarise_all(funs(mean,se))
kable(t(summary.end), caption="Body composition of animals which have completed the study")

summary.end.diff <-
  summary.end %>%
  select(Knockout,Gender,ends_with('mean')) %>%
  group_by(Knockout,Gender) %>%
  gather(key="Variable",value="Measurement",-c(Gender,Knockout)) %>%
  spread(key="Knockout", value="Measurement") %>%
  mutate(Difference = `TRUE`-`FALSE`,
         Difference.pct = Difference/`FALSE`*100)

kable(summary.end.diff, caption="Effects at the end of the study")
```


# Session Information
```{r sessionInfo}
sessionInfo()
```

