---
title: "Tissue Weights for Muscle Tsc1 Knockout Mice"
author: "Dave Bridges and Erin Stephenson"
date: "2019-02-08"
output:
  pdf_document:
    highlight: tango
    keep_tex: yes
    number_sections: yes
    toc: yes
  html_document:
    highlight: tango
    keep_md: yes
    number_sections: yes
    toc: yes
---

```{r global_options, include=FALSE}
library(knitr)
#figures makde will go to directory called figures, will make them as both png and pdf files 
opts_chunk$set(fig.path='figures/',
               echo=FALSE, warning=FALSE, message=FALSE,dev=c('png','pdf'))
options(scipen = 2, digits = 3)
# set echo and message to TRUE if you want to display code blocks and code output respectively

knitr::knit_hooks$set(inline = function(x) {
  knitr:::format_sci(x, 'md')
})


superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
  arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
  length = length, ...)

  
se <- function(x) sd(x, na.rm=T)/sqrt(length(x))

#load these packages, nearly always needed
library(tidyr)
library(dplyr)

# sets maize and blue color scheme
color.scheme <- c('#00274c', '#ffcb05')
```

# Purpose

To determine tissue weights at sacrifice for fat pads and muscle tissues

# Experimental Details

At sacrifice, after a 16h fast data were entered and collected in the raw data sheet


```{r data-input}
library(readr) #loads the readr package
filename <- 'Sacrifice Data.csv' #make this a separate line, you can use any variable you want

#this loads whatever the file is into a dataframe called exp.data if it exists
if(filename %in% dir()){
  exp.data <- read_csv(filename,
                       col_types = cols(Genotype= col_factor(levels=NULL),
                                        Sex = col_factor(levels=NULL)))
}
```

These data can be found in **`r getwd()`** in a file named **`r ifelse(filename %in% dir(), filename, "no file found")`**.  This script was most recently updated on **`r date()`**.

# Analysis

```{r summary-data}
mean.na <- function(x) mean(x, na.rm=T)
shapiro.p <- function(x) shapiro.test(x)$p.value

summary.data <-
  exp.data %>%
  select(Genotype,Sex,GonadalWAT,InguinalWAT,Quadriceps,TricepsSurae,Heart) %>%
  mutate(Knockout = Genotype == 'fl/fl; Tg/+') %>%
  group_by(Sex,Knockout) %>%
  summarize_if(is.numeric, funs(mean.na,se,length)) %>%
  filter(!(is.na(Knockout))) %>%
  filter(!(is.na(Sex)))


kable(summary.data %>% select(ends_with('mean.na')))
kable(summary.data %>% select(ends_with('length')))

```

## Fat Pad Weights

```{r wat-weights, fig.cap="Weights of Fat Pads at Sacrifice"}
library(RColorBrewer)
genotype.colors <- brewer.pal(2,"Set2")

library(ggplot2)
library(gridExtra)
gWAT <- ggplot(summary.data,
       aes(y=GonadalWAT_mean.na,
           ymax=GonadalWAT_mean.na+GonadalWAT_se,
           ymin=GonadalWAT_mean.na-GonadalWAT_se,
           x=Sex,
           group=Knockout,
           fill=Knockout)) +
  geom_bar(stat='identity',position='dodge',width=0.75) +
  geom_errorbar(position=position_dodge(width=0.75),width=0.5) +
  labs(y="Weight (g)",
          title="Gonadal WAT") +
  scale_fill_manual(values=genotype.colors,labels=c("Wild-Type","Knockout")) +
        theme(legend.position = c(0.25, 0.9),
        legend.background = element_rect(fill=alpha(0.1)),
        legend.text=element_text(size=14),
        axis.text=element_text(size=14),
        axis.title=element_text(size=14),
        legend.title=element_blank())

iWAT <- ggplot(summary.data,
       aes(y=InguinalWAT_mean.na,
           ymax=InguinalWAT_mean.na+InguinalWAT_se,
           ymin=InguinalWAT_mean.na-InguinalWAT_se,
           x=Sex,
           group=Knockout,
           fill=Knockout)) +
  geom_bar(stat='identity',position='dodge',width=0.75) +
  geom_errorbar(position=position_dodge(width=0.75),width=0.5) +
  labs(y="Weight (g)",
          title="Inguinal WAT") +
  scale_fill_manual(values=genotype.colors,labels=c("Wild-Type","Knockout")) +
        theme(legend.position = c(0.25, 0.9),
        legend.background = element_rect(fill=alpha(0.1)),
        legend.text=element_text(size=14),
        axis.text=element_text(size=14),
        axis.title=element_text(size=14),
        legend.title=element_blank())

grid.arrange(gWAT,iWAT,ncol=2)
```

## Muscle Weights

```{r muscle-weights, fig.cap="Weights of Muscle Depots at Sacrifice"}
library(RColorBrewer)
genotype.colors <- brewer.pal(2,"Set2")

library(ggplot2)
library(gridExtra)
Heart <- ggplot(summary.data,
       aes(y=Heart_mean.na,
           ymax=Heart_mean.na+Heart_se,
           ymin=Heart_mean.na-Heart_se,
           x=Sex,
           group=Knockout,
           fill=Knockout)) +
  geom_bar(stat='identity',position='dodge',width=0.75) +
  geom_errorbar(position=position_dodge(width=0.75),width=0.5) +
  labs(y="Weight (g)",
          title="Heart") +
  scale_fill_manual(values=genotype.colors,labels=c("Wild-Type","Knockout")) +
        theme(legend.position = c(0.25, 0.9),
        legend.background = element_rect(fill=alpha(0.1)),
        legend.text=element_text(size=14),
        axis.text=element_text(size=14),
        axis.title=element_text(size=14),
        legend.title=element_blank())

Quad <- ggplot(summary.data,
       aes(y=Quadriceps_mean.na,
           ymax=Quadriceps_mean.na+Quadriceps_se,
           ymin=Quadriceps_mean.na-Quadriceps_se,
           x=Sex,
           group=Knockout,
           fill=Knockout)) +
  geom_bar(stat='identity',position='dodge',width=0.75) +
  geom_errorbar(position=position_dodge(width=0.75),width=0.5) +
  labs(y="Weight (g)",
          title="Quadriceps") +
  scale_fill_manual(values=genotype.colors,labels=c("Wild-Type","Knockout")) +
        theme(legend.position = c(0.25, 0.9),
        legend.background = element_rect(fill=alpha(0.1)),
        legend.text=element_text(size=14),
        axis.text=element_text(size=14),
        axis.title=element_text(size=14),
        legend.title=element_blank())

TS <- ggplot(summary.data,
       aes(y=TricepsSurae_mean.na,
           ymax=TricepsSurae_mean.na+TricepsSurae_se,
           ymin=TricepsSurae_mean.na-TricepsSurae_se,
           x=Sex,
           group=Knockout,
           fill=Knockout)) +
  geom_bar(stat='identity',position='dodge',width=0.75) +
  geom_errorbar(position=position_dodge(width=0.75),width=0.5) +
  labs(y="Weight (g)",
          title="Triceps Surae") +
  scale_fill_manual(values=genotype.colors,labels=c("Wild-Type","Knockout")) +
        theme(legend.position = c(0.25, 0.9),
        legend.background = element_rect(fill=alpha(0.1)),
        legend.text=element_text(size=14),
        axis.text=element_text(size=14),
        axis.title=element_text(size=14),
        legend.title=element_blank())

grid.arrange(Heart,Quad, TS,ncol=3)

```

# Session Information

```{r session-information, echo=T}
sessionInfo()
```