---
title: "ITT for mTSC1-HFD mice"
author: "Erin Stephenson and Dave Bridges"
date: "March 8, 2016"
output:
  html_document:
    highlight: tango
    keep_md: yes
    number_sections: yes
    toc: yes
  pdf_document:
    highlight: tango
    keep_tex: yes
    number_sections: yes
    toc: yes
---

```{r global_options, include=FALSE}
library(knitr)
#figures makde will go to directory called figures, will make them as both png and pdf files 
opts_chunk$set(fig.path='figures/',
               echo=TRUE, warning=FALSE, message=FALSE,dev=c('png','pdf'))
options(scipen = 2, digits = 3)
# set echo and message to TRUE if you want to display code blocks and code output respectively

knitr::knit_hooks$set(inline = function(x) {
  knitr:::format_sci(x, 'md')
})


superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
  arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
  length = length, ...)

  
se <- function(x) sd(x, na.rm=T)/sqrt(length(x))

#load these packages, nearly always needed
library(tidyr)
library(dplyr)
```

# Data Entry

```{r data-entry}
library(rjson)
library(RCurl)
#download data from experiment 1417
itt_url <- 'http://bridgeslab.sph.umich.edu/mousedb/api/v1/data/?format=json&username=davebridges&api_key=fce3fde2a9a2e5dc9e04c20aad90120a621c50b3&limit=0&experiment=1417'
itt.json <- fromJSON(getURL(itt_url))
library(plyr)
#parse json data into data frame
itt.data <-  rbind(ldply(lapply(itt.json[[2]],function(x) do.call(c,unlist(x,r=F))),as.data.frame))
#figure out which mice are mTSC1-KO

#set fl/fl;Tg/+ as the reference level
itt.data$genotype <- relevel(as.factor(itt.data$animal.Genotype), ref='fl/fl; Tg/+')
itt.data$Sex <- relevel(as.factor(itt.data$animal.Gender), ref='F')
itt.data <-
  itt.data %>%
  mutate(Knockout = factor(if_else(genotype!='fl/fl; Tg/+', "Wild-Type", "Knockout")))

itt.glucose <- itt.data[itt.data$assay.assay=='Plasma Glucose',]                   
#separate comma separated list into values
s <- strsplit(as.character(itt.glucose$values), split = ",")
#paste values into new colunms
itt.glucose <- cbind(itt.glucose, as.data.frame(rbind.fill.matrix(lapply(s, t))))
#convert all values to numbers, put in new columns
itt.glucose$t0 <- as.numeric(as.character(itt.glucose$`1`))
itt.glucose$t15 <- as.numeric(as.character(itt.glucose$`2`))
itt.glucose$t30 <- as.numeric(as.character(itt.glucose$`3`))
itt.glucose$t45 <- as.numeric(as.character(itt.glucose$`4`))
itt.glucose$t60 <- as.numeric(as.character(itt.glucose$`5`))
itt.glucose$t75 <- as.numeric(as.character(itt.glucose$`6`))
itt.glucose$t90 <- as.numeric(as.character(itt.glucose$`7`))
itt.glucose$t105 <- as.numeric(as.character(itt.glucose$`8`))
itt.glucose$t120 <- as.numeric(as.character(itt.glucose$`9`))

color.palette <- c("blue4","dodgerblue", "deeppink4","pink2")

library(readr)
write_csv(itt.glucose, "ITT Data.csv")
```

```{r glucose-levels, dev=c('png','pdf')}
itt.glucose <- read_csv("ITT Data.csv")
library(dplyr)
#generate averages/se/n/shapiro test for each time point
se <- function(x) sd(x, na.rm=T)/sqrt(length(x))
itt.summary.means <- 
  itt.glucose %>%
  dplyr::group_by(Sex,Knockout) %>%
  dplyr::summarize(
    t0 = mean(t0, na.rm=T),
    t15 = mean(t15, na.rm=T),
    t30 = mean(t30, na.rm=T),
    t45 = mean(t45, na.rm=T),
    t60 = mean(t60, na.rm=T),
    t75 = mean(t75, na.rm=T),
    t90 = mean(t90, na.rm=T),
    t105 = mean(t105, na.rm=T),
    t120 = mean(t120, na.rm=T))

itt.summary.se <- 
  itt.glucose %>%
  dplyr::group_by(Sex,Knockout) %>%
  dplyr::summarize(
    t0 = se(t0),
    t15 = se(t15),
    t30 = se(t30),
    t45 = se(t45),
    t60 = se(t60),
    t75 = se(t75),
    t90 = se(t90),
    t105 = se(t105),
    t120 = se(t120))

times <- seq(0,120,15)
ymax <- max(itt.summary.means[,3:dim(itt.summary.means)[2]]) + max(itt.summary.se[,3:dim(itt.summary.se)[2]])

par(mfrow=c(1,2))
plot <- plot(times,
     as.numeric(itt.summary.means[itt.summary.means$Knockout=='Wild-Type'&itt.summary.means$Sex=="M",3:dim(itt.summary.means)[2]]), 
     las=1, ylim=c(0,ymax), 
     ylab="Blood Glucose (mg/dL)", 
     xlab="Time After Insulin Injection (min)", 
     type="l",
     col=color.palette[1],
     main="Male Mice")
#add ko line
lines(times, as.numeric(itt.summary.means[itt.summary.means$Knockout=='Knockout'&itt.summary.means$Sex=="M",3:dim(itt.summary.means)[2]]), col=color.palette[2])
#add error bars
superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
length = length, ...)

#add error bars for wt and ko
superpose.eb(times,
             as.numeric(itt.summary.means[itt.summary.means$Knockout=='Wild-Type'&itt.summary.means$Sex=="M",3:dim(itt.summary.means)[2]]),
             as.numeric(itt.summary.se[itt.summary.se$Knockout=='Wild-Type'&itt.summary.se$Sex=="M",3:dim(itt.summary.se)[2]]),
             col=color.palette[1])
superpose.eb(times,
             as.numeric(itt.summary.means[itt.summary.means$Knockout=='Knockout'&itt.summary.means$Sex=="M",3:dim(itt.summary.means)[2]]),
             as.numeric(itt.summary.se[itt.summary.se$Knockout=='Knockout'&itt.summary.se$Sex=="M",3:dim(itt.summary.se)[2]]),
             col=color.palette[2])
#add legend
legend("bottomleft",c("Wild-Type",'Knockout'), col=color.palette[1:2], lty=1, bty='n')


plot <- plot(times,
     as.numeric(itt.summary.means[itt.summary.means$Knockout=='Wild-Type'&itt.summary.means$Sex=="F",3:dim(itt.summary.means)[2]]), 
     las=1, ylim=c(0,ymax), 
     ylab="Blood Glucose (mg/dL)", 
     xlab="Time After Insulin Injection (min)", 
     type="l",
     col=color.palette[3],
     main="Female Mice")
#add ko line
lines(times, as.numeric(itt.summary.means[itt.summary.means$Knockout=='Knockout'&itt.summary.means$Sex=="F",3:dim(itt.summary.means)[2]]), col=color.palette[4])
#add error bars for wt and ko
superpose.eb(times,
             as.numeric(itt.summary.means[itt.summary.means$Knockout=='Wild-Type'&itt.summary.means$Sex=="F",3:dim(itt.summary.means)[2]]),
             as.numeric(itt.summary.se[itt.summary.se$Knockout=='Wild-Type'&itt.summary.se$Sex=="F",3:dim(itt.summary.se)[2]]),
             col=color.palette[3])
superpose.eb(times,
             as.numeric(itt.summary.means[itt.summary.means$Knockout=='Knockout'&itt.summary.means$Sex=="F",3:dim(itt.summary.means)[2]]),
             as.numeric(itt.summary.se[itt.summary.se$Knockout=='Knockout'&itt.summary.se$Sex=="F",3:dim(itt.summary.se)[2]]),
             col=color.palette[4])
#add legend
legend("bottomleft",c("Wild-Type",'Knockout'), col=color.palette[3:4], lty=1, bty='n')

```

```{r normalized}

#normalized means and errors to 100%
itt.summary.means.norm <- sweep(itt.summary.means[3:dim(itt.summary.means)[2]], 1, itt.summary.means$t0, "/")*100
itt.summary.se.norm <- sweep(itt.summary.se[3:dim(itt.summary.se)[2]], 1, itt.summary.means$t0, "/")*100

par(mfrow=c(1,2))
plot <- plot(times,
     as.numeric(itt.summary.means.norm[3,1:dim(itt.summary.means.norm)[2]]), 
     las=1, ylim=c(0,ymax), 
     ylab="Blood Glucose (% of fasted)", 
     xlab="Time After Insulin Injection (min)", 
     type="l",
     col=color.palette[1],
     main="Male Mice")
#add ko line
lines(times, as.numeric(itt.summary.means.norm[4,1:dim(itt.summary.means.norm)[2]]), col=color.palette[2])

#add error bars for wt and ko
superpose.eb(times,
             as.numeric(itt.summary.means.norm[3,1:dim(itt.summary.means.norm)[2]]),
             as.numeric(itt.summary.se.norm[3,1:dim(itt.summary.se.norm)[2]]),
             col=color.palette[1])
superpose.eb(times,
             as.numeric(itt.summary.means.norm[4,1:dim(itt.summary.means.norm)[2]]),
             as.numeric(itt.summary.se.norm[4,1:dim(itt.summary.se.norm)[2]]),
             col=color.palette[2])
#add legend
legend("bottomleft",c("Wild-Type",'Knockout'), col=color.palette[1:2], lty=1, bty='n')


plot <- plot(times,
     as.numeric(itt.summary.means.norm[1,1:dim(itt.summary.means.norm)[2]]), 
     las=1, ylim=c(0,ymax), 
     ylab="Blood Glucose (% of fasted)", 
     xlab="Time After Insulin Injection (min)", 
     type="l",
     col=color.palette[3],
     main="Female Mice")
#add ko line
lines(times, as.numeric(itt.summary.means.norm[2,1:dim(itt.summary.means.norm)[2]]), col=color.palette[4])
#add error bars for wt and ko
superpose.eb(times,
             as.numeric(itt.summary.means.norm[1,1:dim(itt.summary.means.norm)[2]]),
             as.numeric(itt.summary.se.norm[1,1:dim(itt.summary.se.norm)[2]]),
             col=color.palette[3])
superpose.eb(times,
             as.numeric(itt.summary.means.norm[2,1:dim(itt.summary.means.norm)[2]]),
             as.numeric(itt.summary.se.norm[2,1:dim(itt.summary.se.norm)[2]]),
             col=color.palette[4])
#add legend
legend("bottomleft",c("Wild-Type",'Knockout'), col=color.palette[3:4], lty=1, bty='n')
```


# Statistics

```{r statistics}
#mixed linear model
library(lme4)
library(reshape2)
melted.data <- melt(itt.glucose[,c('animal.id','genotype',paste("t", seq(0,120,15), sep=""))],
                    id.vars=c('genotype','animal.id'), value.name="glucose", variable.name='time')

long.data <- 
  itt.glucose %>%
  select(Knockout,Sex,animal.MouseID,t0:t120) %>%
  group_by(Knockout,Sex, animal.MouseID) %>%
  gather(key=Time, value=Glucose, -Knockout, -Sex, -animal.MouseID)

fit.null <- lmer(Glucose~Time+(1|animal.MouseID), data=long.data, REML=F)
fit.null.sex <- lmer(Glucose~Time+Sex+(1|animal.MouseID), data=long.data, REML=F) # was significantly improved
fit.null.sex.int <- lmer(Glucose~Time+Sex+Sex:Time+(1|animal.MouseID), 
                         data=long.data, REML=F) -> fit.null #was also significantly improved, used as null model
fit.ko <- lmer(Glucose~Time+Sex+Sex:Time+Knockout+(1|animal.MouseID), 
                         data=long.data, REML=F) #was significant relative to null

fit.ko.int.time <- lmer(Glucose~Time+Sex+Sex:Time+Knockout+Knockout:Time+(1|animal.MouseID), 
                         data=long.data, REML=F) #was significant relative to fit.ko

fit.ko.int.time <- lmer(Glucose~Time+Sex+Sex:Time+Knockout+Knockout:Time+(1|animal.MouseID), 
                         data=long.data, REML=F) #was significant relative to fit.ko

fit.ko.sex <- lmer(Glucose~Time + Knockout + Knockout:Time * Sex +(1|animal.MouseID), 
                         data=long.data, REML=F) #suggests significant modifying effect of sex over model without sex
fit.ko.sex.null <- lmer(Glucose~Time + Knockout + Knockout:Time + Sex + (1|animal.MouseID), 
                         data=long.data, REML=F) 
library(broom)
anova(fit.ko.sex, fit.ko.sex.null) %>% tidy() %>% kable(caption="Model testing for moderating effect of sex")

#calculated AUC
auc.data <-
  itt.glucose %>%
  select(Knockout,Sex,animal.MouseID,t0:t120) %>%
  mutate(AUC = t0+t15+t30+t45+t60+t75+t90+t105+t120) 

auc.summary <-
  auc.data %>%
  group_by(Sex,Knockout) %>%
  summarize_at('AUC', .funs=funs(mean(.,na.rm=T), se, shapiro.test(.)$p.value))

kable(auc.summary, caption="Summary statistics for AUC")

auc.f <- auc.summary %>% filter(Sex=="F") %>% pull(mean)
auc.m <- auc.summary %>% filter(Sex=="M") %>% pull(mean)
library(car)
```

We calculated the AUC and found a `r -(auc.f[1]-auc.f[2])/auc.f[2]*100` improvement in insulin sensitivity in females (**p=`r wilcox.test(AUC~Knockout, data=auc.data %>% filter(Sex=="F"))$p.value`** from a Mann-Whitney test) and a `r -(auc.m[1]-auc.m[2])/auc.m[2]*100` improvement in male mice (levene Test p=`r leveneTest(AUC~Knockout, data=auc.data %>% filter(Sex=="M"))$"Pr(>F)"[1]`; Welch's *t* test **p=`r t.test(AUC~Knockout, var.equal=F, data=auc.data %>% filter(Sex=="F"))$p.value`**).

## Session Information

```{r session-info}
sessionInfo()
```
