---
title: "Tissue Weights for Muscle Tsc1 Knockout Mice on HFD"
author: "Dave Bridges and Erin Stephenson"
date: "2019-03-01"
output:
  pdf_document:
    highlight: tango
    keep_tex: yes
    number_sections: yes
    toc: yes
  html_document:
    highlight: tango
    keep_md: yes
    number_sections: yes
    toc: yes
---

```{r global_options, include=FALSE}
library(knitr)
#figures makde will go to directory called figures, will make them as both png and pdf files 
opts_chunk$set(fig.path='figures/',
               echo=FALSE, warning=FALSE, message=FALSE,dev=c('png','pdf'))
options(scipen = 2, digits = 3)
# set echo and message to TRUE if you want to display code blocks and code output respectively

knitr::knit_hooks$set(inline = function(x) {
  knitr:::format_sci(x, 'md')
})


superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
  arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
  length = length, ...)

  
se <- function(x) sd(x, na.rm=T)/sqrt(length(x))

#load these packages, nearly always needed
library(tidyr)
library(dplyr)

# sets maize and blue color scheme
color.scheme <- c('#00274c', '#ffcb05')
```

# Purpose

To determine tissue weights at sacrifice for fat pads and muscle tissues

# Experimental Details

At sacrifice, after a 16h fast data were entered and collected in the raw data sheet


```{r data-input}
library(readxl) #loads the readr package
filename <- 'HFD Sacrifice Data.xlsx' #make this a separate line, you can use any variable you want

#this loads whatever the file is into a dataframe called exp.data if it exists
exp.data <- read_excel(filename,
                     sheet = 'RawData')
```

These data can be found in **`r getwd()`** in a file named **`r ifelse(filename %in% dir(), filename, "no file found")`**.  This script was most recently updated on **`r date()`**.

# Analysis

```{r summary-data}
mean.na <- function(x) mean(x, na.rm=T)
shapiro.p <- function(x) shapiro.test(x)$p.value

summary.data <-
  exp.data %>%
  select(Genotype,Sex,GonadalWAT,InguinalWAT,Quadriceps,TricepsSurae,Heart) %>%
  mutate(Knockout = Genotype == 'fl/fl; Tg/+') %>%
    filter(!(is.na(Genotype))) %>%
  mutate(Sex = relevel(as.factor(Sex), ref="Male")) %>%
  group_by(Sex,Knockout) %>%
  summarize_if(is.numeric, funs(mean.na,se,length,shapiro.p)) %>%
  filter(!(is.na(Knockout))) %>%
  filter(!(is.na(Sex)))


kable(summary.data %>% select(ends_with('mean.na')))
kable(summary.data %>% select(ends_with('length')))

```

## Fat Pad Weights

```{r wat-weights, fig.cap="Weights of Fat Pads at Sacrifice"}
library(RColorBrewer)
genotype.colors <- brewer.pal(2,"Set2")

library(ggplot2)
library(gridExtra)
gWAT <- ggplot(summary.data,
       aes(y=GonadalWAT_mean.na,
           ymax=GonadalWAT_mean.na+GonadalWAT_se,
           ymin=GonadalWAT_mean.na-GonadalWAT_se,
           x=Sex,
           group=Knockout,
           fill=Knockout)) +
  geom_bar(stat='identity',position='dodge',width=0.75) +
  geom_errorbar(position=position_dodge(width=0.75),width=0.5) +
  labs(y="Weight (g)",
          title="Gonadal WAT") +
  scale_fill_manual(values=genotype.colors,labels=c("Wild-Type","Knockout")) +
        theme(legend.position = c(0.25, 0.9),
        legend.background = element_rect(fill=alpha(0.1)),
        legend.text=element_text(size=14),
        axis.text=element_text(size=14),
        axis.title=element_text(size=14),
        legend.title=element_blank())

iWAT <- ggplot(summary.data,
       aes(y=InguinalWAT_mean.na,
           ymax=InguinalWAT_mean.na+InguinalWAT_se,
           ymin=InguinalWAT_mean.na-InguinalWAT_se,
           x=Sex,
           group=Knockout,
           fill=Knockout)) +
  geom_bar(stat='identity',position='dodge',width=0.75) +
  geom_errorbar(position=position_dodge(width=0.75),width=0.5) +
  labs(y="Weight (g)",
          title="Inguinal WAT") +
  scale_fill_manual(values=genotype.colors,labels=c("Wild-Type","Knockout")) +
        theme(legend.position = c(0.25, 0.9),
        legend.background = element_rect(fill=alpha(0.1)),
        legend.text=element_text(size=14),
        axis.text=element_text(size=14),
        axis.title=element_text(size=14),
        legend.title=element_blank())

grid.arrange(gWAT,iWAT,ncol=2)

#using base R graphics
par(mfrow=c(1,2))
plot.data <- 
  summary.data %>%
  select(Sex, Knockout, InguinalWAT_mean.na) %>%
  spread(value=InguinalWAT_mean.na, key=Sex) 

plot.data.se <- 
  summary.data %>%
  select(Knockout, InguinalWAT_se) %>%
  spread(value=InguinalWAT_se, key=Sex)   

ymax <- max(plot.data[2:3]) + max(plot.data.se[2:3])

plot <- barplot(as.matrix(plot.data[2:3]),
                las=1,
                names.arg=c("Male","Female"),
                ylab="Weight (g)",
                beside=T,
                main="Inguinal WAT",
                ylim=c(0,ymax))

superpose.eb(plot,
             as.matrix(plot.data[2:3]),
             as.matrix(plot.data.se[2:3]))

plot.data <- 
  summary.data %>%
  select(Sex, Knockout, GonadalWAT_mean.na) %>%
  spread(value=GonadalWAT_mean.na, key=Sex) 

plot.data.se <- 
  summary.data %>%
  select(Knockout, GonadalWAT_se) %>%
  spread(value=GonadalWAT_se, key=Sex)   

ymax <- max(plot.data[2:3]) + max(plot.data.se[2:3])

plot <- barplot(as.matrix(plot.data[2:3]),
                las=1,
                names.arg=c("Wild-Type","Knockout"),
                ylab="Weight (g)",
                beside=T,
                main="Gonadal WAT",
                ylim=c(0,ymax))

superpose.eb(plot,
             as.matrix(plot.data[2:3]),
             as.matrix(plot.data.se[2:3]))

```

For the male mice, the fat pads were reduced in weight:

```{r wat-differences}
library(forcats)
summary.data %>%
  select(Sex, Knockout, GonadalWAT_mean.na) %>%
  mutate(Knockout = factor(Knockout, labels=c("Wild-Type","Knockout"))) %>%
  spread(value=GonadalWAT_mean.na, key=Knockout) %>%
  mutate(Difference = `Wild-Type`-Knockout,
         Pct.Difference = Difference/`Wild-Type`*100) %>%
  kable(caption="Changes in Gonadal Fat Pad Weights")


summary.data %>%
  select(Sex, Knockout, InguinalWAT_mean.na) %>%
  mutate(Knockout = factor(Knockout, labels=c("Wild-Type","Knockout"))) %>%
  spread(value=InguinalWAT_mean.na, key=Knockout) %>%
  mutate(Difference = `Wild-Type`-Knockout,
         Pct.Difference = Difference/`Wild-Type`*100) %>%
  kable(caption="Changes in Inguinal Fat Pad Weights")
  
```

```{r male-stats}
summary.data %>%
  select(Sex,
         Knockout,
         InguinalWAT_shapiro.p,
         GonadalWAT_shapiro.p) %>%
  kable(caption="Shapiro-Wilk Tests for each group")

library(car)
stats.ing <- 
  exp.data %>%
  mutate(Knockout = factor(if_else(Genotype != 'fl/fl; Tg/+', "Wild-Type", "Knockout"))) %>%
  group_by(Sex) %>%
  filter(!(is.na(Genotype))) %>%
  summarize(
    Levene = leveneTest(InguinalWAT~Knockout)$"Pr(>F)"[1],
    Mann.Whitney = wilcox.test(InguinalWAT~Knockout)$p.value,
    Welch = t.test(InguinalWAT~Knockout, var.equal=F)$p.value,
    Student = t.test(InguinalWAT~Knockout, var.equal=T)$p.value)

stats.gon <- 
  exp.data %>%
  mutate(Knockout = factor(if_else(Genotype != 'fl/fl; Tg/+', "Wild-Type", "Knockout"))) %>%
  group_by(Sex) %>%
  filter(!(is.na(Genotype))) %>%
  summarize(
    Levene = leveneTest(GonadalWAT~Knockout)$"Pr(>F)"[1],
    Mann.Whitney = wilcox.test(GonadalWAT~Knockout)$p.value,
    Welch = t.test(GonadalWAT~Knockout, var.equal=F)$p.value,
    Student = t.test(GonadalWAT~Knockout, var.equal=T)$p.value)

  kable(stats.ing, caption="Pairwise tests for effects of knockout on Inguinal WAT weights.")
    kable(stats.gon, caption="Pairwise tests for effects of knockout on Gonadal WAT weights.")
```

For the male mice, normality can be assumed, but not equal variance, so a Welch's t-test is used, which had a p-value of `r stats.ing[2,'Welch']` for inguinal WAT and `r stats.gon[2,'Welch']` for gonadal WAT.

For the female mice, normality cannot be assumed so a Mann-Whitney test is used, which had a p-value of `r stats.ing[1,'Mann.Whitney']` for inguinal WAT and `r stats.gon[1,'Mann.Whitney']` for gonadal WAT.

## Muscle Weights

```{r muscle-weights, fig.cap="Weights of Muscle Depots at Sacrifice"}
library(RColorBrewer)
genotype.colors <- brewer.pal(2,"Set2")

library(ggplot2)
library(gridExtra)
Heart <- ggplot(summary.data,
       aes(y=Heart_mean.na,
           ymax=Heart_mean.na+Heart_se,
           ymin=Heart_mean.na-Heart_se,
           x=Sex,
           group=Knockout,
           fill=Knockout)) +
  geom_bar(stat='identity',position='dodge',width=0.75) +
  geom_errorbar(position=position_dodge(width=0.75),width=0.5) +
  labs(y="Weight (g)",
          title="Heart") +
  scale_fill_manual(values=genotype.colors,labels=c("Wild-Type","Knockout")) +
        theme(legend.position = c(0.25, 0.9),
        legend.background = element_rect(fill=alpha(0.1)),
        legend.text=element_text(size=14),
        axis.text=element_text(size=14),
        axis.title=element_text(size=14),
        legend.title=element_blank())

Quad <- ggplot(summary.data,
       aes(y=Quadriceps_mean.na,
           ymax=Quadriceps_mean.na+Quadriceps_se,
           ymin=Quadriceps_mean.na-Quadriceps_se,
           x=Sex,
           group=Knockout,
           fill=Knockout)) +
  geom_bar(stat='identity',position='dodge',width=0.75) +
  geom_errorbar(position=position_dodge(width=0.75),width=0.5) +
  labs(y="Weight (g)",
          title="Quadriceps") +
  scale_fill_manual(values=genotype.colors,labels=c("Wild-Type","Knockout")) +
        theme(legend.position = c(0.25, 0.9),
        legend.background = element_rect(fill=alpha(0.1)),
        legend.text=element_text(size=14),
        axis.text=element_text(size=14),
        axis.title=element_text(size=14),
        legend.title=element_blank())

TS <- ggplot(summary.data,
       aes(y=TricepsSurae_mean.na,
           ymax=TricepsSurae_mean.na+TricepsSurae_se,
           ymin=TricepsSurae_mean.na-TricepsSurae_se,
           x=Sex,
           group=Knockout,
           fill=Knockout)) +
  geom_bar(stat='identity',position='dodge',width=0.75) +
  geom_errorbar(position=position_dodge(width=0.75),width=0.5) +
  labs(y="Weight (g)",
          title="Triceps Surae") +
  scale_fill_manual(values=genotype.colors,labels=c("Wild-Type","Knockout")) +
        theme(legend.position = c(0.25, 0.9),
        legend.background = element_rect(fill=alpha(0.1)),
        legend.text=element_text(size=14),
        axis.text=element_text(size=14),
        axis.title=element_text(size=14),
        legend.title=element_blank())

grid.arrange(Heart,Quad, TS,ncol=3)

```

# Session Information

```{r session-information, echo=T}
sessionInfo()
```