DESeq Analysis
========================================================

```{r setup}
input_directory <- "../data/raw/htseq-counts"
sample_mapping_file <- '../data/raw/sample_mapping.csv'
```

This file uses the counts table in `r input_directory` from `r sample_mapping_file`.  This script was most recently run on `r date()`

```{r data-entry, echo=FALSE, message=FALSE}
sampleFiles <- list.files(input_directory)
mappings <- read.csv(sample_mapping_file)
sampleTable <- data.frame(sampleName = mappings$File,
  fileName = mappings$File,
  condition = mappings$Genotype)
library(DESeq2)

dds.binary <- DESeqDataSetFromHTSeqCount(sampleTable = sampleTable,
  directory = input_directory,
  design= ~ condition)
dds.binary$condition <- relevel(dds.binary$condition,"Wild-Type")

dds.binary <- DESeq(dds.binary)
res.binary <- results(dds.binary)
resOrdered.binary <- res.binary[order(res.binary$padj),]
```

```{r annotation, echo=FALSE}
library(biomaRt)
#download gene name annotation data for results sets.
ensembl <- useMart("ensembl",dataset = 'mmusculus_gene_ensembl')
transcriptInfo <- getBM(
    attributes=c('external_gene_id','ensembl_gene_id'), 
    values = 'protein_coding',
    filter = 'biotype',
    mart=ensembl)
annotated_data <- merge(as.data.frame(resOrdered.binary), transcriptInfo, by.x='row.names', by.y='ensembl_gene_id')
ordered_annotated_data <- annotated_data[order(annotated_data$log2FoldChange),]

output_file_binary <- "../data/processed/Binary DESeq Results.csv"
write.csv(as.data.frame(annotated_data),
  file=output_file_binary)

summary.table <- table(ordered_annotated_data$padj<0.05, ordered_annotated_data$log2FoldChange>0)
rownames(summary.table) <- paste("Significant", rownames(summary.table))
colnames(summary.table) <- paste("Upregulated", colnames(summary.table))
```

The calculated data was exported to `r output_file_binary`.  There was a total of **`r table(ordered_annotated_data$padj<0.05)[2]`** significantly different genes out of `r table(ordered_annotated_data$padj<0.05)[1] + table(ordered_annotated_data$padj<0.05)[2]` that met our testing criteria.  These are broken down into up and downregulated genes below:

```{r summary-table, echo=FALSE, results='asis', message=FALSE}
library(xtable)
print(xtable(summary.table, caption="Number of Genes Up/Downregulated and Significant or Not"),type='html')
```

Diagnostic Plots
----------------

```{r ma-plots, echo=FALSE, dev=c('png','pdf')}
plotMA(res.binary, main="DESeq2", ylim=c(-4,4), title="MA Plot for Binary Comparison")
```

```{r heatmap, echo=FALSE, dev=c('png','pdf')}
library("RColorBrewer")
library("gplots")
select <- rownames(subset(as.data.frame(res.binary), padj<0.05))
hmcol <- colorRampPalette(brewer.pal(9, "GnBu"))(100)
heatmap.2(counts(dds.binary,normalized=TRUE)[select,], col = hmcol,
Rowv = F, Colv = T, scale="row", trace="none", margin=c(10,6))

rld.binary <- rlog(dds.binary)
vsd.binary <- varianceStabilizingTransformation(dds.binary)
rlogMat.binary <- assay(rld.binary)
vstMat.binary <- assay(vsd.binary)

#rld transformed heatmap
heatmap.2(assay(rld.binary)[select,], col = hmcol,
  Rowv = FALSE, Colv = FALSE, scale="row",
  dendrogram="none", trace="row", margin=c(10, 6))

#vsd stabilized heatmap
heatmap.2(assay(vsd.binary)[select,], col = hmcol,
  Rowv = FALSE, Colv = FALSE, scale="row",
  dendrogram="none", trace="none", margin=c(10, 6))

```

Plotting Estimated Dispersions
-------------------------------
```{r dispersion-plot, echo=FALSE, dev=c('png','pdf')}
plotDispEsts(dds.binary)
```


Principal Component Analysis
------------------------------

```{r pca-plot, echo=FALSE, dev=c('png','pdf')}
plotPCA(rld.binary, intgroup=c("condition"))
```


Session Information
-------------------

```{r session-info}
sessionInfo()
```